---
title: An Analysis of Texas High School Academic Competition Results, Part 5 - Miscellaneous
slug: analysis-texas-high-school-academics-5-miscellaneous
date: "2018-05-20"
categories: []
tags:
 - r
 - analysis
 - texas
 - academics
banner: "img/analysis-texas-high-school-academics/miscellaneous-banner.png"
---
```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  # echo = TRUE,
  echo = FALSE,
  cache = TRUE,
  # include = FALSE,
  fig.align = "center",
  # results = "asis",
  fig.width = 7,
  fig.height = 7,
  # out.width = 7,
  # out.height = 7,
  warning = FALSE,
  message = FALSE
)

```

```{r load, include = FALSE}
library("tidyverse")
library("rlang")
library("teplot")
library("teproj")
library("tetidy")
load(file.path("static", "data", "uil-v02-setup.RData"))
config$path_schools_uil <-
  "https://raw.githubusercontent.com/tonyelhabr/uil-v02/master/data/schools-uil.csv"
schools_uil <-
  config$path_schools_uil %>%
  # teproj::import_path_cleanly()
  readr::read_csv()
config$path_persons_uil <-
  "https://raw.githubusercontent.com/tonyelhabr/uil-v02/master/data/persons-uil.csv"
persons_uil <-
  config$path_persons_uil %>%
  # teproj::import_path_cleanly()
  readr::read_csv()
```
There's a lot to analyze with the Texas high school academic UIL data set.
Maybe I find it more interesting than others due to my personal experiences with
these competitions.

Now, after examining some of the biggest topics associated
with this data--including competitions, individuals, and schools--in a broad manner,
there are
some other things that don't necessarily fall into these categories that 
I think are worth investigating.

## Siblings


Let's look at the performance of siblings. Maybe this topic only came to mind for
me because I have brothers, on of who is my twin, but I think anyone can find
something interesting on this matter.

### Sibling Participation

So, let's start with something easy--which siblings competed together the most?

<include>
As in the analysis of individuals and schools, I'll include myself and my school in the tables and figures
where applicable.
</include>

```{r siblings_funcs, include = FALSE, dependson = c("load")}
add_name_first_pair_col <-
  function(data = NULL) {
    data %>% 
      mutate(name_first_pair = paste0(name_first, " & ", name_first_sibling))
  }

get_siblings_by_at <-
  function(data = NULL,
           min = 2,
           max = Inf,
           cols_grp = NULL,
           prettify = TRUE) {

    data_proc <-
      data %>%
      group_by(!!!syms(cols_grp)) %>%
      mutate(rnk_max = rank(name_last, ties.method = "max")) %>%
      ungroup()

    data_filt <-
      data_proc %>%
      filter(rnk_max >= min, rnk_max <= max)

    ret <-
      data_filt %>%
      inner_join(data_filt,
                by = cols_grp,
                suffix = c("", "_sibling")) %>%
      filter(name_first != name_first_sibling)
    
    # ret <-
    #   ret %>% 
    #   distinct(!!!syms(cols_grp), .keep_all = TRUE)
    
    if(prettify) {
      ret <-
        ret %>% 
        add_name_first_pair_col() %>% 
        select(
          name_first_pair,
          name,
          name_first,
          name_last,
          name_first_sibling,
          everything()) %>%
        arrange(name_last, name_first, school)
      
      cols_drop <-
        setdiff(c(names(data), "name_first_pair", "name_first_sibling"), names(ret))
      
      ret <-
        ret %>%
        select(-one_of(cols_drop))
    }
      ret
  }

```


```{r siblings, include = FALSE, dependson = c("load")}
siblings <-
  persons_uil %>%
  get_siblings_by_at(
    cols_grp = c("name_last",
                 "school",
                 "city",
                 "year",
                 "conf",
                 "complvl",
                 "comp",
                 "complvl_num")
  )
```

```{r siblings_n, include = FALSE, dependson = c("load", "siblings")}
siblings_n <-
  siblings %>% 
  group_by(name_last, name_first_pair) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  arrange(n) %>% 
  distinct(name_last, n, .keep_all = TRUE) %>%
  tetidy::rank_arrange_at("n")

```

```{r html_siblings_n, echo = FALSE, include = TRUE, dependson = c("load", "siblings")}
html_siblings_n <-
  siblings_n %>%
  create_kable_filt_at(
    config = config,
    col_rgx = "name_last",
  )
html_siblings_n
```
Admittedly, I am a bit disappointed to find that my twin brother and I are not at
the very top of this list. Nonetheless, we are fairly near the top, so I can take
some satisfaction in that. [^twins]

[^twins]:
I don't explicitly try to filter for twins only, but it's reasonable to believe
that many, if not most, are twins.

I should note that the scraped data does not distinguish siblings, so 
I had to define criteria to do so.
To be specific, the table above enforces the criteria that two people
have the same last name, school, and city, and that they compete
in the exact same competition--that is, a competition occurring in a given year and 
being of a same competition type and same competition level (as well as
the same conference and competition area, if applicable). The numbers
are inflated when not enforcing the criteria that the two people must have competed
in the same competition type and level (nor conference and competition area), and even more so
when throwing out the criteria for same year.


```{r siblings_lax, include = FALSE, dependson = c("load"), eval = FALSE}
siblings_lax <-
  persons_uil %>%
  get_siblings_by_at(
    cols_grp = c("name_last", "school", "city")
  )

siblings_lax_n <-
  siblings_lax %>%
  group_by(school, name_last, name_first_pair) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  distinct(school, name_last, n, .keep_all = TRUE) %>%
  tetidy::rank_arrange_at("n")
siblings_lax_n
```

```{r html_siblings_lax_n, echo = FALSE, include = TRUE, dependson = c("load", "siblings_lax_n"), eval = FALSE}
html_siblings_lax_n <-
  siblings_lax_n %>%
  mutate_at(vars(n), funs(scales::comma)) %>% 
  create_kable_filt_at(
    config = config,
    col_rgx = "name_last"
  )
html_siblings_lax_n
```

### Sibling Performance

Participation in competitions is one thing, but what about sibling performance?
Let's use the same metric used elsewhere for ranking performance--percent rank of
scores summed across all records (`prnk_sum`)--and see which sibling pairs
show up among the top.

```{r siblings_prnk, include = FALSE, dependson = c("siblings")}
siblings_prnk <-
  siblings %>%
  group_by(name_last, name_first_pair) %>%
  summarise_at(vars(matches("^prnk|^rnk|^n_")), funs(sum)) %>% 
  ungroup() %>%
  distinct(name_last, prnk_sum, .keep_all = TRUE) %>%
  gather(metric_sibling, value_sibling, matches("_sibling$")) %>% 
  mutate_at(vars(metric_sibling), funs(str_replace(., "_sibling", ""))) %>% 
  gather(metric, value, matches("^prnk|^rnk|^n_")) %>% 
  filter(metric == metric_sibling) %>% 
  mutate(sum = value + value_sibling) %>% 
  select(-matches("metric_sibling|value")) %>% 
  spread(metric, sum) %>% 
  tetidy::rank_arrange_at("prnk")
siblings_prnk
```

```{r html_siblings_prnk, echo = FALSE, include = TRUE, dependson = c("load", "siblings_prnk")}
html_siblings_prnk <-
  siblings_prnk %>%
  mutate_if(is.numeric, funs(round(., 2))) %>% 
  mutate_at(vars(starts_with("n_")), funs(scales::comma)) %>% 
  create_kable_filt_at(
    config = config,
    col_rgx = "name_last"
  )
html_siblings_prnk
```

It looks like the pairs at the top of these rankings based on score are fairly similar to the list
of pairs competing most frequently. (This is not too surprising
given that my choice of metric of ranking is based on a summed value that "rewards"
volume of participation rather than per-competition performance.) Again, my twin brother and I
appear near the top.

## My High School

Even though I highlighted my high school ("CLEMENS") in my examination of schools and looked
at individual scores elsewhere, I did not
look at other individuals that have gone to my school. Perhaps it is a bit egotistical,
but I am interested in knowing how I compare with others that have attended my school (either
before, with, or after me).

```{r persons_n_filt, include = FALSE, dependson = c("load")}
persons_n_filt <-
  persons_uil %>%
  add_stats_cols_by_at(cols_grp = c("name", "school", "city")) %>% 
  filter(str_detect(school, config$rgx_school_filt)) %>%
  tetidy::rank_arrange_at("prnk_sum") %>% 
  select(-school, -city, -n_state_sum)
persons_n_filt
```

```{r html_persons_n_filt, echo = FALSE, include = TRUE, dependson = c("load", "persons_n_filt")}
html_persons_n_filt <-
  persons_n_filt %>% 
  mutate_if(is.numeric, funs(round(., 2))) %>% 
  create_kable_filt_at(
    config = config,
    col_rgx = "name"
  )
html_persons_n_filt
```

Alas, although my twin brother and I did not rank among the very top of the siblings
by participation and performance, we do appear among the top 
when evaluating only people from my high school. In my opinion, the sample size isn't so small
that this achievement is trivial. 

## Wrap-Up

I think all I've done here is more investigation of my personal performance, 
so I'll spare the reader any more of my egotistical exporation.
And, with that said, I think this is a good point to bring an end to my
investigation of Texas high school academic UIL competitions.
