---
title: An Analysis of Texas High School Academic Competition Results, Part 2 - Competitions
slug: analysis-texas-high-school-academics-2-competitions
date: "2018-05-20"
categories: []
tags:
 - r
 - analysis
 - texas
 - academics
banner: "img/analysis-texas-high-school-academics/viz_n_bycompcomplvlconf.png"
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  # echo = TRUE,
  echo = FALSE,
  cache = TRUE,
  # include = FALSE,
  fig.align = "center",
  # results = "asis",
  fig.width = 7,
  fig.height = 7,
  # out.width = 7,
  # out.height = 7,
  warning = FALSE,
  message = FALSE
)

```

```{r load, include = FALSE}
library("tidyverse")
library("rlang")
library("teplot")
library("teproj")
library("tetidy")
load(file.path("static", "data", "uil-v02-setup.RData"))
config$path_schools_uil <-
  "https://raw.githubusercontent.com/tonyelhabr/uil-v02/master/data/schools-uil.csv"
schools_uil <-
  config$path_schools_uil %>%
  # teproj::import_path_cleanly()
  readr::read_csv()
config$path_persons_uil <-
  "https://raw.githubusercontent.com/tonyelhabr/uil-v02/master/data/persons-uil.csv"
persons_uil <-
  config$path_persons_uil %>%
  # teproj::import_path_cleanly()
  readr::read_csv()

summarise_n_bygrps <-
  function(data = NULL,
           cols_grp1 = NULL,
           cols_grp2 = NULL,
           cols_grp3 = NULL) {
    ret <-
      data %>%
      group_by(!!!syms(cols_grp1)) %>%
      summarise(n = n()) %>%
      ungroup()

    if (!is.null(cols_grp2)) {
      ret <-
        ret %>%
        group_by(!!!syms(cols_grp2)) %>%
        summarise(n = mean(n)) %>%
        ungroup()
    }

    if (!is.null(cols_grp3)) {
      ret <-
        ret %>%
        group_by(!!!syms(cols_grp3)) %>%
        summarise(n = mean(n)) %>%
        ungroup()
    }
    ret %>%
      tetidy::rank_arrange_at("n")
  }

summarise_n_bycomplvl <-
  function(data = NULL, complvls = NULL) {
    data %>%
      filter(complvl %in% complvls) %>%
      distinct(year, complvl_num, school) %>%
      summarise_n_bygrps(
        cols_grp1 = c("year", "complvl_num"),
        cols_grp2 = c("complvl_num")
      ) %>%
      mutate(complvl = complvls)
  }

summarise_n_bycomp <-
  function(data = NULL) {
    data %>%
      summarise_n_bygrps(cols_grp1 = c("year", "comp"),
                         cols_grp2 = c("comp"))
  }

summarise_n_bycompcomplvl <-
  function(data = NULL) {
    data %>%
      summarise_n_bygrps(
        cols_grp1 = c("year", "complvl", "comp"),
        cols_grp2 = c("complvl", "comp")
      )
  }

summarise_n_bycompcomplvlconf <-
  function(data = NULL) {
    data %>%
      summarise_n_bygrps(
        cols_grp1 = c("year", "complvl", "comp", "conf"),
        cols_grp2 = c("complvl", "comp", "conf")
      ) %>%
      group_by(complvl, comp) %>%
      tetidy::rank_arrange_at("n") %>%
      ungroup() %>%
      arrange(comp, complvl, rnk)
  }

scale_color_set1 <- function() {
  scale_color_brewer(palette = "Set1")
}

scale_fill_set1 <- function() {
  scale_fill_brewer(palette = "Set1")
}
```

## Competition Participation

Some of the first questions that might come to mind are those regarding the 
number of schools
in each level of competition (District, Region, and State) and each 
conference classification level (1A, 2A, ... 6A).

```{r summ_n_bycomplvl, include = FALSE, dependson = c("load")}
summ_n_bycomplvl <-
  bind_rows(
    schools_uil %>%
      summarise_n_bycomplvl("District"),
    schools_uil %>%
      summarise_n_bycomplvl("Region"),
    schools_uil %>%
      mutate(complvl = "Conference") %>%
      select(-complvl_num) %>%
      rename(complvl_num = conf) %>%
      summarise_n_bycomplvl("Conference")
  )
```

```{r viz_n_bycomplvl, echo = FALSE, include = TRUE, fig.height = 7, fig.width = 7, dependson = c("load", "summ_n_bycomplvl")}
viz_n_bycomplvl <-
  summ_n_bycomplvl %>%
  ggplot(aes(x = complvl_num, y = n)) +
  geom_point() +
  geom_hline(
    data = summ_n_bycomplvl %>%
      group_by(complvl) %>%
      summarise(n_mean = mean(n)),
    aes(yintercept = n_mean),
    color = "black",
    linetype = "dashed",
    size = 2
  ) +
  teplot::scale_y_pretty_comma() +
  facet_wrap (~ complvl, scales = "free") +
  labs(
    title = "Average Annual Count of Schools by Different UIL Groups",
    caption =
      paste0(
        config$viz_footer,
        str_wrap(
          paste0(
            "Dashed line: average for each grouping. ",
            "Note that these groupings are not subsets of the same \"parent\" group. ",
            "(Rather, they are different groupings of all schools .",
            "District and Region are references to competition levels,",
            " while conference is a classification based on school size.) ",
            "State is not shown because there is no \"subgroup\" of it."
          ),
        100
        )
      )
  ) +
  labs(x = "Group \"Number\"", y = "Aveaage Annual Count") +
  teplot::theme_te_facet() +
  theme(legend.position = "none")
viz_n_bycomplvl
teproj::export_ext_png(
  viz_n_bycomplvl,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 7,
  width = 7
)
```

It seems fair to say that the distribution of schools among Districts, Regions, and Conferences
is relatively even. [^state_grouping] [^competition_groupings] This is to be expected since the UIL presumably
tries to divide schools evenly among each grouping (to the extent possible) in
order to stimulate fair competition.

[^state_grouping]:
The State competition level is not shown because it there is no "sub-grouping" of State
(like there is 1, 2, 3, ... for each of the groupings shown).

[^competition_groupings]:
As a technical note, Districts, Regions, and Conferences are not really all of the same "type"
of data. Nevertheless, these different "groupings" each stratify the sample population in some manner.

With some context regarding the number of schools in each competition level and in conference,
let's now consider the number of distinct individual competitors and schools for each
competition type (i.e. Calculator, Computer Science, etc.)


```{r summ_n_bycomp, include = FALSE, dependson = c("load")}
summ_n_bycomp <-
  bind_rows(
    schools_uil %>%
      summarise_n_bycomp() %>%
      mutate(entity = "Schools"),
    persons_uil %>%
      summarise_n_bycomp() %>%
      mutate(entity = "Competitors")
  )

```

```{r visualize_n_bycomp_common_at, include = FALSE, dependson = c("load")}
visualize_n_bycomp_common_at <-
  function(data = NULL,
           x = "comp",
           col_grp = NULL) {
    data %>%
      # inner_join(comp_icons, by = "comp") %>%
      ggplot(aes_string(x = x, y = "n")) +
      geom_point(aes(color = comp), size = 5) +
      # geom_icon(aes(image = icon, color = icon), size = 0.1) +
      geom_hline(
        data =
          data %>%
          group_by(!!sym(col_grp)) %>%
          mutate(n_mean = mean(n)),
        aes(yintercept = n_mean),
        color = "black",
        linetype = "dashed",
        size = 1
      ) +
      scale_color_set1() +
      guides(color = guide_legend(override.aes = list(size = 5))) +
      scale_x_discrete(labels = scales::wrap_format(10)) +
      teplot::scale_y_pretty_comma() +
      teplot::labs_xy_null() +
      teplot::theme_te_facet()
  }

```
```{r comp_icons, include = FALSE, eval = FALSE}
# comp_icons <-
#   tibble::tibble(
#     comp = config$comps_info$comp_name,
#     icon = c("calculator", "laptop", "stats-bars", "wand", "flask")
#   )
```

```{r viz_n_bycomp, echo = FALSE, include = TRUE, fig.height = 7, fig.width = 7, dependson = c("load", "summ_n_bycomp")}
viz_n_bycomp <-
  summ_n_bycomp %>%
  visualize_n_bycomp_common_at(col_grp = "entity") +
  facet_wrap(~ entity, scales = "free") +
  theme(axis.text.x = element_blank()) +
  labs(
    title = "Count of Entities for Each Competition Type",
    caption = config$viz_footer
  )
viz_n_bycomp

teproj::export_ext_png(
  viz_n_bycomp,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 7,
  width = 7
)
```

Science stands out when evaluating participation by competition type alone.

But what about when considering competition level as well?

```{r summ_n_bycompcomplvl, include = FALSE, dependson = c("load")}
summ_n_bycompcomplvl <-
  bind_rows(
    schools_uil %>%
      summarise_n_bycompcomplvl() %>%
      mutate(entity = "Schools"),
    persons_uil %>%
      summarise_n_bycompcomplvl() %>%
      mutate(entity = "Competitors")
  ) %>%
  mutate(complvl_entity = paste(complvl, entity, sep = ", "))
```


```{r viz_n_bycompcomplvl, echo = FALSE, include = TRUE, fig.height = 7, fig.width = 7, dependson = c("load", "summ_n_bycompcomplvl")}
viz_n_bycompcomplvl <-
  summ_n_bycompcomplvl %>%
  visualize_n_bycomp_common_at(col_grp = "complvl_entity") +
  facet_wrap( ~ complvl_entity, scales = "free", ncol = 2) +
  theme(axis.text.x = element_blank()) +
  labs(
    title = "Count of Entities for Each Competition Type and Level",
    caption = config$viz_footer
  )
viz_n_bycompcomplvl

teproj::export_ext_png(
  viz_n_bycompcomplvl,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 7,
  width = 7
)
```

Science seems to prevail again.

Now, what about when also considering conferences?

```{r summ_n_bycompcomplvlconf, include = FALSE, dependson = c("load")}
summ_n_bycompcomplvlconf <-
  bind_rows(
    schools_uil %>%
      summarise_n_bycompcomplvlconf() %>%
      mutate(entity = "Schools"),
    persons_uil %>%
      summarise_n_bycompcomplvlconf() %>%
      mutate(entity = "Competitors")
  ) %>%
  mutate(complvl_comp = paste(complvl, comp, sep = ", "))
summ_n_bycompcomplvlconf
```

```{r viz_n_bycompcomplvlconf, echo = FALSE, include = TRUE, fig.height = 7, fig.width = 7, dependson = c("load", "summ_n_bycompcomplvlconf")}
viz_n_bycompcomplvlconf <-
  summ_n_bycompcomplvlconf %>%
  filter(entity == "Competitors") %>%
  mutate_at(vars(conf), funs(factor)) %>%
  visualize_n_bycomp_common_at(x = "conf", col_grp = "complvl_comp") +
  facet_grid(complvl ~ comp, scales = "free", labeller = label_wrap_gen(width = 12)) +
  labs(
    title = "Count of Competition for Each Competition Type,\nLevel, and Conference",
    caption = config$viz_footer,
    x = "Conference"
  ) +
  theme(legend.position = "none")
viz_n_bycompcomplvlconf

teproj::export_ext_png(
  viz_n_bycompcomplvlconf,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 7,
  width = 7
)
```

Once again, Science seems to be the most popular.
So, in all, science appears to be the answer to the numerous
variations of the "Which competition type/level has the most ...?" question.

### Competition Scores

With an understanding of the participation in the UIL competitions,
let's now consider the scores. What does the distribution of scores
for each competition type look like?

```{r persons_stats_bycomp, include = FALSE, dependson = c("load")}
summarise_stats_score_by_at <-
  function(data = NULL,
           col = "score",
           cols_grp = NULL) {
    tetidy::summarise_stats_by_at(data = data,
                                  col = col,
                                  cols_grp = cols_grp)
  }

persons_stats_bycomp <-
  persons_uil %>%
  summarise_stats_score_by_at(cols_grp = "comp")
persons_stats_bycomp
```

```{r viz_persons_stats_bycomp, echo = FALSE, include = TRUE, fig.width = 7, fig.height = 10, dependson = c("load", "persons_stats_bycomp")}
n_comps <- length(config$comps_info_list)
q05_min <- min(persons_stats_bycomp$q05)
q95_max <- max(persons_stats_bycomp$q95)

lab_title_stats_bycomp_prefix <-
  "Distribution of Scores for Each Competition Type"
lab_caption_stats_bycomp <-
  paste0(
    config$viz_footer,
    str_wrap("Dashed line: mean. Solid lines: z-score = -1 and 1.", 60)
  )

viz_persons_stats_bycomp <-
  persons_uil %>%
  inner_join(persons_stats_bycomp, by = "comp") %>%
  ggplot(aes(x = score)) +
  geom_histogram(aes(fill = comp), bins = 30) +
  geom_density() +
  scale_fill_set1() +
  geom_vline(
    aes(xintercept = mean),
    color = "black",
    linetype = "solid",
    size = 1
  ) +
  geom_vline(aes(xintercept = zn1), color = "black", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = zp1), color = "black", linetype = "dashed", size = 1) +
  guides(fill = FALSE) +
  scale_x_continuous(limits = c(q05_min, q95_max)) +
  # scale_y_continuous(breaks = scales::pretty_breaks(), labels = scales::comma) +
  teplot::scale_y_pretty_comma() +
  # scale_y_continuous(breaks = function(x) round(x, 2)) +
  facet_wrap(
    ~ comp,
    scales = "free",
    nrow = n_comps,
    labeller = label_wrap_gen(width = 12),
    strip.position = "right"
  ) +
  teplot::labs_xy_null() +
  labs(title = lab_title_stats_bycomp_prefix,
       caption = lab_caption_stats_bycomp) +
  teplot::theme_te_facet() +
  theme(legend.position = "none")
viz_persons_stats_bycomp

teproj::export_ext_png(
  viz_persons_stats_bycomp,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 10,
  width = 7
)
```

Take what you will from the above visual, but it is interesting to see which
competition types have wider distributions. This could imply a number of different things:

+ A competition type with a wider distribution (relative to the others) may
be more likely to attract "newbies".

+ Or, such a distribution may simply be headed by "experts".

+ Or, perhaps, it might simply imply that the test was not well made (or "fair"),
resulting in a wide variety of scores. (On the other hand, one might simply
see a right- or left-skewed distribution if the test is not robust.)

To try to understand the score distributions better, let's break them
down by year, competition level, and conference.

```{r persons_stats_bycompx, include = FALSE, dependson = c("load")}
persons_stats_bycompyear <-
  persons_uil %>%
  summarise_stats_score_by_at(cols_grp = c("comp", "year"))

persons_stats_bycompcomplvl <-
  persons_uil %>%
  summarise_stats_score_by_at(cols_grp = c("comp", "complvl"))

persons_stats_bycompconf <-
  persons_uil %>%
  summarise_stats_score_by_at(cols_grp = c("comp", "conf"))
```

```{r viz_persons_stats_bycompx, include = FALSE, results = "hide", dependson = c("load", "persons_stats_bycompx")}
visualize_persons_stats_bycompx_at <-
  function(data = NULL, x = NULL) {
    # Using facet_wrap() so strip.position can be specified.
    data %>%
      ggplot(aes_string(x = x, y = "mean")) +
      geom_pointrange(aes(
        ymin = zn1,
        ymax = zp1,
        color = comp
      )) +
      scale_color_set1() +
      guides(color = FALSE) +
      facet_wrap(
        ~ comp,
        scales = "free",
        nrow = n_comps,
        labeller = label_wrap_gen(width = 12),
        strip.position = "right"
      ) +
      coord_flip() +
      teplot::labs_xy_null() +
      teplot::theme_te_facet() +
      theme(legend.position = "none")
  }

year_min <- min(persons_uil$year)
year_max <- max(persons_uil$year)
year_labs <- seq.int(year_min, year_max, by = 4L)

lab_caption_stats_bycompx <-
  gsub("Dashed line", "Point", lab_caption_stats_bycomp)

viz_persons_stats_bycompyear <-
  persons_stats_bycompyear %>%
  visualize_persons_stats_bycompx_at(x = "year") +
  scale_x_continuous(breaks = year_labs) +
  # scale_x_continuous(breaks = scales::pretty_breaks()) +
  labs(title = lab_title_stats_bycomp_prefix,
       subtitle = "By Year",
       caption = lab_caption_stats_bycompx)
viz_persons_stats_bycompyear

viz_persons_stats_bycompcomplvl <-
  persons_stats_bycompcomplvl %>%
  visualize_persons_stats_bycompx_at(x = "complvl") +
  labs(title = lab_title_stats_bycomp_prefix,
       subtitle = "By Competition Level",
       caption = lab_caption_stats_bycompx)
viz_persons_stats_bycompcomplvl

viz_persons_stats_bycompconf <-
  persons_stats_bycompconf %>%
  visualize_persons_stats_bycompx_at(x = "conf") +
  labs(title = lab_title_stats_bycomp_prefix,
       subtitle = "By Conference",
       caption = lab_caption_stats_bycompx)
teproj::export_ext_png(
  viz_persons_stats_bycompconf,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 7,
  width = 7
)
```

```{r viz_persons_stats_bycomp_grid, echo = FALSE, include = TRUE, fig.height = 10, fig.width = 7, dependson = c("load", "viz_persons_stats_bycompx")}
viz_persons_stats_bycomp_grid <-
  gridExtra::arrangeGrob(
    viz_persons_stats_bycompyear + theme(strip.text = element_blank()),
    viz_persons_stats_bycompcomplvl + theme(strip.text = element_blank()) + labs(title = "", caption = " \n "),
    viz_persons_stats_bycompconf + labs(title = "", caption = " \n "),
    ncol = 3
  )
gridExtra::grid.arrange(viz_persons_stats_bycomp_grid)

teproj::export_ext_png(
  # viz_persons_stats_bycomp_grid,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 10,
  width = 7,
  viz_persons_stats_bycomp_grid
)
```

One might observe the following from the above plot:

+ It appears from the by-year breakdown that there is no discernible trends among score distributions
across years. The lack of pattern in score distributions
over time is a strong indication that the tests have about the same difficulty every year.
This implies that it would not be unreasonable to compare raw test scores
for a given competition type when comparing individuals
and schools across different years.

+ It is evident from the competition type visual that score range distributions
shift upwards with increasing competition level, as would be expected with
more "superior" competitors advancing to higher levels of competition.

+ From the conference visual, it is apparent that scores generally increase
with increasing conference number, which, itself, corresponds with increasing student body size.
This could be indicative of a number of factors, including:

    - Larger schools have more "superior" individual competitors, perhaps due
    to having a larger "talent pool" of students.
        
    - In an effort to compete in as many different competition types as
    possible and/or reach the minimum number of individual competitors needed to qualify
    the school as a team (echo.g. at least 3 individuals), 
    scmaller schools push individuals to participate in competitions
    even if they are not really prepared to do so.
    (This might be considered a form of [selection bias](https://en.wikipedia.org/wiki/Selection_bias).)

So maybe there isn't some broad, temporal trend to the scores.
But is there a trend in scores among the competition levels (for a given competition type)?
One would think that, assuming that test difficulty is constant
across competition levels, 
there would be an aggregate increase in scores
with increasing competition level (because only the top scoring individuals and schools advance).

```{r persons_stats_bycompcomplvl_diffs, include = FALSE, dependson = c("load", "persons_stats_bycompx")}
persons_stats_bycompcomplvl_diffs <-
  persons_stats_bycompcomplvl %>%
  group_by(comp) %>%
  # mutate_at(vars(mean, median, sd), funs(diff_pct = diff(c(NA, .)) / .))
  mutate_at(
    vars(mean, sd),
    funs(
    diff1_pct = (diff(c(NA, .), lag = 1) / lag(., 1)),
    diff2_pct = (diff(c(NA, NA, .), lag = 2) / lag(., 2))
  )) %>%
  mutate_at(
    vars(complvl),
    funs(
      complvl_diff1 = paste0(lag(.), " to ", .),
      complvl_diff2 = paste0(lag(., 2), " to ", .)
    )
  )

persons_stats_bycompcomplvl_diffs_tidy <-
  persons_stats_bycompcomplvl_diffs %>%
  filter(!is.na(mean_diff1_pct) | !is.na(mean_diff2_pct)) %>%
  gather(complvl_diff, complvl_diff_lab, complvl_diff1, complvl_diff2) %>%
  filter(!is.na(complvl_diff_lab)) %>%
  gather(metric, value, mean_diff1_pct, mean_diff2_pct) %>%
  filter(!is.na(value)) %>%
  filter(!(complvl_diff == "complvl_diff1" & metric == "mean_diff2_pct")) %>%
  filter(!(complvl_diff == "complvl_diff2" & metric == "mean_diff1_pct")) %>%
  ungroup() %>%
  arrange(comp)
```

```{r viz_persons_stats_bycompcomplvl_diffs, echo = FALSE, include = TRUE, fig.height = 7, fig.width = 7, dependson = c("load", "persons_stats_bycompcomplvl_diffs")}
lab_title_persons_stats_bycompcomplvl_diffs <-
  str_wrap("% Change in Average Scores for Each Competition Type, By Competition Level", 60)
viz_persons_stats_bycompcomplvl_diffs <-
  persons_stats_bycompcomplvl_diffs_tidy  %>%
  mutate_at(
    vars(complvl_diff_lab),
    funs(factor),
    levels = c("District to Region", "Region to State", "District to State")
  ) %>%
  ggplot(aes(x = complvl_diff_lab, y = value)) +
  geom_col(aes(fill = comp, alpha = complvl_diff_lab)) +
  scale_fill_set1() +
  scale_alpha_discrete(range = c(0.4, 1)) +
  guides(fill = FALSE, size = FALSE) +
  teplot::scale_y_pretty_percent() +
  facet_grid( ~ comp, scales = "free", labeller = label_wrap_gen(width = 12)) +
  teplot::theme_te_facet_dx() +
  theme(axis.text.x = element_blank()) +
  teplot::labs_xy_null() +
  labs(title = lab_title_persons_stats_bycompcomplvl_diffs,
       caption = config$viz_footer)
viz_persons_stats_bycompcomplvl_diffs

teproj::export_ext_png(
  viz_persons_stats_bycompcomplvl_diffs,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 7,
  width = 7
)
```

In fact, it does appear that score distributions skew higher with increasing levels
of competition. If this were not true, then I would suspect that test are made to be
more difficult at each competition stage. [^increase]

[^increase]:
Actually, it might be true that there _is_ some increase
in test difficulty with advancing competitions, but such an adjustment, if it is truly made,
does not quite offset the skill level of the advancing competitors.

Notably, it appears that Number Sense demonstrates the largest
"jumps" in aggregate scores with increasing competition levels. Having competed
in this competition before, I do not find this all too surprising.
More than any other competition type, those who succeed Number Sense seem to
rely on natural abilities (as opposed to training) to beat out the competition.
(I would go so far as to say that for some individuals, this natural ability is "savant-like".)
Consequently, with increasing competition level, it is more likely that these
"superior" competitors stand out, and, as observed here, cause the
the scoring distributions to skew higher.

So there __does__ seem to be a trend of higher scores with higher competition level,
but has __that__ trend changed over time? [^trend]

[^trend]:
Recall that there is no apparent trend when not distinguishing by competition level.

```{r persons_stats_bycompyear_terms, include = FALSE, dependson = c("load", "persons_stats_bycompx")}
persons_stats_bycompyear_diffs <-
  persons_stats_bycompyear %>%
  mutate_at(vars(mean, median, sd), funs(diff_pct = diff(c(NA, .)) / .))

# fmla_fit <- formula(mean_diff_pct ~ mean + comp)
fmla_fit <- formula(mean_diff_pct ~ mean + year)
persons_stats_bycompyear_terms <-
  persons_stats_bycompyear_diffs %>%
  # group_by(year) %>%
  group_by(comp) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~lm(fmla_fit, data = .x))) %>%
  mutate(terms = purrr::map(fit, ~broom::tidy(.x))) %>%
  unnest(terms, .drop = TRUE) %>%
  arrange(comp, p.value)
persons_stats_bycompyear_terms
```

```{r viz_persons_stats_bycompyear_diffs, echo = FALSE, include = TRUE, fig.height = 7, fig.width = 7, dependson = c("load", "persons_stats_bycompyear_terms")}
lab_title_persons_stats_bycompyear_diffs <-
  str_wrap("Year-to-Year % Change in Average Scores for Each Competition Type, By Competition Level", 60)
lab_caption_persons_stats_bycompyear_diffs <-
  paste0(
    config$viz_footer,
    str_wrap(
      "Dashed line: Competition-specific average % change across all year-to-year intervals.",
    100
    )
  )
viz_persons_stats_bycompyear_diffs <-
  persons_stats_bycompyear_diffs %>%
  ggplot(aes(x = year, y = mean_diff_pct)) +
  geom_line(aes(color = comp), size = 2) +
  scale_color_set1() +
  geom_smooth(method = "lm", se = FALSE, color = "black", linetype = "dashed", size = 1) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  teplot::scale_y_pretty_percent() +
  guides(color = FALSE) +
  # geom_smooth(se = FALSE, color = "black", size = 1) +
  facet_wrap( ~ comp,
              scales = "fixed",
              ncol = 1,
              labeller = label_wrap_gen(width = 12),
              strip.position = "right") +
  teplot::theme_te_facet() +
  theme(legend.position = "none") +
  teplot::labs_xy_null() +
  labs(title = lab_title_persons_stats_bycompyear_diffs,
       subtitle = "By Year",
       caption = lab_caption_persons_stats_bycompyear_diffs)
viz_persons_stats_bycompyear_diffs

teproj::export_ext_png(
  viz_persons_stats_bycompyear_diffs,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 7,
  width = 7
)
```

As with the raw scores, there does not appear to be any noticeable
trend in the change in level of difficulty of tests between each competition level
over time. (Aside from a steep drop-off in the Science scores in 2014, there's really
nothing too interesting about this plot.)
Along with the previous temporal visual where a trend with competition level is not distinguished,
this plot is a strong indication that the tests (and, most likely,
the skills of the competitors) have not changed over time.

Such visual inferences are supported quantitatively.
Observe the 
[p-values](https://en.wikipedia.org/wiki/P-value) of the variables 
for [linear regression](https://en.wikipedia.org/wiki/Linear_regression)
models fit for each competition type,
where the average year-to-year score difference (as a percentage) is estimated given
average individual competitor score and year. Taking the customary p-value threshold of 0.05
as the level of significance (where one may deduce that the
model is not predictive if the p.value is greater than the threshold value), the year term
is shown to be insignificant.

$$
mean \space year \space diff \space pct = intercept + mean * \beta_{1} + year * \beta_{2}
$$

```{r html_persons_stats_bycompyear_terms, echo = FALSE, include = TRUE, dependson = c("load", "persons_stats_bycompyear_terms")}
html_persons_stats_bycompyear_terms <-
  persons_stats_bycompyear_terms %>%
  # filter(term == "year") %>%
  select(comp, term, p.value) %>%
  mutate_if(is.numeric, funs(round(., 4))) %>%
  spread(term, p.value) %>%
  teproj::create_kable(n_show = Inf)
html_persons_stats_bycompyear_terms
```

There could be a number of confounding factors that explain the relatively even level of
competition over time, including the following.

+ It could be that test makers have purposely adjusted the difficulty of tests
to result in scores that are similar to historical scores for a given competition
and/or competition level.

+ Or, perhaps more likely, the difficulty of tests and the skill level of competitors
have stayed relatively the same over time.

If the latter is true, at the very least
we can say that, despite how the media or popular culture may portray younger
generations, the prowess of the academic "elite" of the current generation of high school
students in Texas has not dropped off.
(Nonetheless, nothing definitive should be deduced about those 
who do not compete in UIL academic competitions and are anywhere outside of Texas.)

## Wrap-up

In this write-up, the participation and scores of Texas high school 
academic UIL competitions are illustrated in a number of different
ways--by competition type, by competition level, by year, and by variations of these.
Next, we'll take a closer look at some more specific questions, focusing on individual competitors.
