---
title: An Analysis of Texas High School Academic Competition Results, Part 1 - Introduction
slug: analysis-texas-high-school-academics-intro
date: "2018-05-20"
categories: []
tags:
 - r
 - analysis
 - texas
 - academics
banner: "img/analysis-texas-high-school-academics/intro-banner.png"
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  # echo = TRUE,
  echo = FALSE,
  cache = TRUE,
  # include = FALSE,
  fig.align = "center",
  # results = "asis",
  fig.width = 7,
  fig.height = 7,
  # out.width = 7,
  # out.height = 7,
  warning = FALSE,
  message = FALSE
)

```

```{r load, include = FALSE}
library("tidyverse")
library("rlang")
library("teplot")
library("teproj")
library("tetidy")
load(file.path("static", "data", "uil-v02-setup.RData"))
config$path_schools_nces_join <-
  "https://raw.githubusercontent.com/tonyelhabr/uil-v02/master/data/schools-nces-join.csv"
schools_nces_join <-
  config$path_schools_nces_join %>%
  # teproj::import_path_cleanly()
  readr::read_csv()
```


<!--<div class="toggle"><button>Show</button>-->

__NOTE:__
<i>
This is part of a series of posts discussing my findings of
[Texas high school academic University Interscholastic Scholarship (UIL) competitions](http://www.uiltexas.org/academics).

To keep this and the other write-ups concise and to focus reader attention on the content,
I have decided not to show the underlying code
(especially that which is used to create the visuals). Nonetheless, the full code
can be viewed on [my GitHub account](https://github.com/tonyelhabr). In the future, I may write some
some kind of addendum to demonstrate some of the programming implementation that I
think is notable in some way.
</i>
<!--</div>-->

## Introduction

After I finished high school in 2012, I thought it would be interesting
to look back and evaluate my
performance in the academic [University Interscholastic League](http://www.uiltexas.org/) (UIL)
competitions that I competed in with historical results.
(To provide some background, most public high schools in Texas are registered in 
the [University Interscholastic League](http://www.uiltexas.org/) (UIL), which 
"exists to provide educational extracurricular academic, athletic, and music contests".
For those familiar with the [National Collegiate Athletic Association](https://www.ncaa.com/) (NCAA),
the UIL serves an analogous role for Texas high school extracurricular activities.)

Aside from my own self-interest in the historical results in these competitions,
I think that this analysis can provide some insight into which schools (and individual students)
are really the most "elite". School-wide and invidividual average scores on state- and 
national- standardized tests (e.g. the [SAT](https://en.wikipedia.org/wiki/SAT))
certainly are the most common measure of academic strength,
but I think rankings by academic competitions may be more indicative.

### About the Data

To make some sense of the my analysis, the reader should be aware of the following details about the data.

+ The competition data was scraped from https://www.hpscience.net/ for all years from 2008 through 2017. [^robots]
The data is not listed in an extremely user-friendly format (in my opinion). Consequently, the "cleaned"
data is imperfect in some ways.

[^robots]:
I checked the site's "robots.txt" file prior to scraping rate limits.

+ The UIL categorizes schools into one of six __"Conferences"__. The conference labels
range from __1A, 2A, ..., 6A__, where the increasing leading digit (i.e. 1, 2, etc.)
generally corresponds to increasing school size.

+ Schools only compete against other schools in their conference.

+ The UIL defines __3 levels of competition__ (in order of "difficulty"): District__, __Region__, and __State__.
These are listed in order of "difficulty". That is, 
Winning a District competitions, results in a Region competition appearance,
and, subsequently, winning a Region competition results in a State competition appearance.
(Keep in mind that schools still only compete against other schools in their same
conference, even as they advance.)

+ The UIL defines __32 total Districts__ in Texas, which are aggregated into __4 Regions__.
(The source of the geo-spatial data is
https://nces.ed.gov/opengis/rest/services/K12_School_Locations/EDGE_GEOCODE_PUBLICSCH_1516/MapServer.)


```{r viz_map_bycomplvl_grid, echo = FALSE, include = TRUE, fig.height = 7, fig.width = 7, dependson = c("load")}
visualize_map_bycomplvl <-
  function(data = NULL) {
    teplot::create_map_base_tx(legend.title = element_text()) +
      geom_polygon(
        data =
          data %>%
          count(county, complvl_num, complvl, sort = TRUE) %>%
          mutate_at(vars(complvl_num), funs(factor)) %>%
          inner_join(
            teplot::get_map_data_county_tx(),
            by = c("county" = "subregion")
          ),
        aes(group = county, fill = complvl_num),
      ) +
      viridis::scale_fill_viridis(# guide = guide_legend(title = complvls, nrow = 4, byrow = TRUE)
        discrete = T,
        option = "D") +
      theme(legend.position = "bottom")
  }

lab_title_map <-
  "Locations of Schools, by Competition Level"

lab_caption_map <-
  str_wrap(
    paste0(
      "Note that some counties may be incorrectly categorized ",
      "due to imperfect data extraction and cleaning. \n\n",
      "Data Source:\n",
      "The geo-spatial data was downloaded from ",
      "https://nces.ed.gov/opengis/rest/services/K12_School_Locations",
      "/EDGE_GEOCODE_PUBLICSCH_1516/MapServer ",
      "and combined with the school information (including districts and regions) ",
      "accompanying the UIL scores ",
      "scraped from https://www.hpscience.net/ via \"fuzzy-joining\"."
      ),
    60)

viz_map_byregion <-
  schools_nces_join %>%
  filter(complvl == "Region") %>%
  visualize_map_bycomplvl() +
  guides(fill = guide_legend(title = "Region", nrow = 1, byrow = FALSE)) +
  theme(plot.caption = element_text(hjust = 0)) +
  labs(title = lab_title_map, caption = lab_caption_map)
viz_map_byregion

viz_map_bydistrict <-
  schools_nces_join %>%
  filter(complvl == "District") %>%
  visualize_map_bycomplvl() +
  guides(fill = guide_legend(title = "District", nrow = 4, byrow = TRUE)) +
  labs(caption = str_replace(config$viz_footer, "\\. Data.*", ""))
viz_map_bydistrict

viz_map_bycomplvl_grid <-
  gridExtra::arrangeGrob(
    viz_map_byregion,
    viz_map_bydistrict,
    nrow = 1
  )

# gridExtra::grid.arrange(viz_map_bycomplvl_grid)

teproj::export_ext_png(
  # viz_map_bycomplvl_grid,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 8,
  width = 10,
  viz_map_bycomplvl_grid
)
```

+ For schools, winning is a "winner-take-all" matter: only the school with the most combined points
among its top handful individual competitors (3 for most competitions) advances.
On the other hand, an individual
may advance even if his school does not win if he places among the top "n".
The value of "n" is dependent on the competition type. [^uil_rules]

+ There are __5 different academic competitions "types"__:
__Calculator Applications, Computer Science, Mathematics, Number Sense, and Science__. [^all_competitions]


[^uil_rules]:
See the [UIL rules](https://www.uiltexas.org/policy/constitution/academics) for more details.

[^all_competitions]:
[There are many more UIL competition types](http://www.uiltexas.org/files/academics/academics-brief-overview.pdf)
than those analyzed here
(including competitions for theater, band, etc.), but these are the ones for academics.

## What's Next

In this series, I investigate the following topics:

+ [The relationships among the competition years, types, levels](/post/analysis-texas-high-school-academics-2-competitions.html)

+ [Individual participation and performance](/post/analysis-texas-high-school-academics-3-individuals.html) (including myself)

+ [School participation and performance](/post/analysis-texas-high-school-academics-4-schools.html) (including my high school)

+ [Miscellaneous topics](/post/analysis-texas-high-school-academics-5-miscellaneous.html), such as sibling performance

