---
title: Visualizing Texas High School SAT Math Scores with Bubble Grids
slug: bubble-grid-texas-high-school
date: "2018-07-22"
categories: []
tags:
 - r
 - ggplot2
 - sf
 - bubble grid
 - texas
 - sat
 - math
banner: "img/analysis-texas-high-school-academics/viz_schools_grid.png"
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  # eval = FALSE
  echo = TRUE,
  # cache = TRUE,
  cache = FALSE,
  include = FALSE,
  fig.align = "center",
  results = "hide",
  fig.width = 6,
  fig.height = 6,
  # out.width = 6,
  # out.height = 6,
  warning = FALSE,
  message = FALSE
)
options(scipen = 1, digits = 2)
```

```{r path_save}
path_save <-
  file.path("static", "data", "tea_bubble_grid.RData")
```


```{r load}
# NOTE TO SELF: Uncomment this and comment out shapefile imports for blog post.
load(file = path_save)
```

Two awesome things inspired this post:

+ [`{ggplot2}`'s version 3.0 release](https://www.tidyverse.org/articles/2018/07/ggplot2-3-0-0/)
on [CRAN](https://cran.r-project.org/), including full support for the 
`{sf}` package and new functions `geom_sf()` and `coord_sf()`, which
make plotting data from shapefiles very straightforward.

+ [Jonas Scholey's blog post](http://jschoeley.github.io/2018/06/30/bubble-grid-map.html)
discussing the use of "bubble grid" maps as an alternative to
[choropleth](https://en.wikipedia.org/wiki/Choropleth_map) maps, which
seem to be used more prevalent.

As Jonas implies, using color as a visual encoding is not always the best option,
a notion with which I strongly agree. [Cartograms](https://en.wikipedia.org/wiki/Cartogram)
try to address the ambiguity of color encoding with distortion of land area/distance,
but I think the result can be difficult to interpret. Bubble grid maps
seem to me to be an interesting alternative that can potentially display
information in a more direct manner.

With that being said, I decided to adapt Jonas' code to visualize
the Texas high school SAT/ACT data that I've looked at in other posts.
To simplify the visual encoding of information,
I'll filter the data down to a single 
statistic---the math test scores for the SAT for the year 2015.
(For other applications, the statistic might be population density, average median household income, etc.)
For the geo-spatial data,
I downloaded the shapefiles
for [schools](https://schoolsdata2-tea-texas.opendata.arcgis.com/datasets/059432fd0dcb4a208974c235e837c94f_0) and
[counties](https://schoolsdata2-tea-texas.opendata.arcgis.com/datasets/c71146b6426248a5a484d8b3c192b9fe_0)
provided by the [Texas Education Agency](https://tea.texas.gov/) (TEA).
Additionally,
I downloaded shapefiles for 
[Texas cities](https://opendata.arcgis.com/datasets/993d420b9f0742b9afa06622d27a37e0_0.geojson) and for
[Texas highways](http://hub.arcgis.com/datasets/e661b4004aee4c939569a563b6bb881a_0),
provided by the [Texas Department of Transportation](https://www.txdot.gov/)
(TxDOT). By plotting the major cities and roadways in the state, the locations
of "sparsely" populated areas should be evident, which can explain
why there doesn't appear to any data in some regions. Finally, I'll
also use the Texas state and county border data provided in the `ggplot2::map_data()`
(which is essentially just serves as a wrapper for extracting
data provided in the `{maps}` package).

```{r config}
# NOTE TO SELF: Set export logicals to FALSE for blog post.
config <-
  list(
    export_data = FALSE,
    dir_data = file.path("static", "data"),
    export_viz = FALSE,
    dir_viz = "figs"
  )
```

```{r packages, include = TRUE}
library("tidyverse")
library("teplot")
library("sf")
```

```{r schools_dl, eval = FALSE}
url_schools <- "https://opendata.arcgis.com/datasets/059432fd0dcb4a208974c235e837c94f_0.geojson"
path_dest_schools <- file.path("data-raw", "schools")
downloader::download(url = url_schools, destfile = path_dest_schools)
```

```{r schools_raw_sf, eval = FALSE}
schools_raw_sf <-
  path_dest %>%
  sf::st_read() %>%
  mutate_if(is.factor, funs(as.character)) %>% 
  janitor::clean_names()
```

```{r schools_sf, eval = FALSE}
schools_sf <-
  schools_raw_sf %>%
  mutate_at(vars(district, campus, county, region, gradegrp), funs(as.integer)) %>%
  rename(
    district_num = district,
    school_num = campus,
    county_num = county,
    region_num = region,
    grade_grp_num = gradegrp
  ) %>%
  rename(
    school = campname,
    district = distname,
    county = cntyname
  ) %>%
  mutate_at(vars(school, district, county, city), funs(toupper)) %>%
  mutate_at(vars(county), funs(str_remove_all(., "\\s+COUNTY"))) %>%
  mutate_at(vars(school), funs(str_remove_all(., "([H]\\s*[S]$)|(\\s+\\&)") %>% str_trim())) %>%
  select(
    school_num,
    school,
    district,
    city,
    county,
    region_num,
    grade_grp = graderange,
    grade_grp_num,
    instr_type,
    magnet,
    geometry
  )
```

```{r path_schools_sf, eval = FALSE}
path_schools_sf <- 
  file.path(config$dir_data, "schools_sf.shp")
```

```{r schools_sf_export, eval = FALSE}
schools_sf %>% 
  st_write(path_schools_sf, delete_dsn = TRUE)
```

```{r schools_sf_import, eval = FALSE}
schools_sf <-
  st_read(path_schools_sf)
```

```{r counties_dl, eval = FALSE}
url_counties <- "https://opendata.arcgis.com/datasets/c71146b6426248a5a484d8b3c192b9fe_0.geojson"
path_dest_counties <- file.path("data-raw", "counties")
downloader::download(url = url_counties, destfile = path_dest_counties)
```

```{r counties_raw_sf, eval = FALSE}
counties_raw_sf <-
  path_dest_counties %>%
  sf::st_read() %>%
  mutate_if(is.factor, funs(as.character)) %>% 
  janitor::clean_names()
```

```{r counties_sf, eval = FALSE}
counties_sf <-
  counties_raw_sf %>%
  mutate(county = fename %>% toupper()) %>%
  select(county, geometry)
counties_sf
```

```{r path_counties_sf, eval = FALSE}
path_counties_sf <- 
  file.path(config$dir_data, "counties_sf.shp")
```

```{r counties_sf_export, eval = FALSE}
counties_sf %>% 
  st_write(path_counties_sf, delete_dsn = TRUE)
```

```{r counties_sf_import, eval = FALSE}
counties_sf <-
  st_read(path_counties_sf)
```

```{r cities_dl, eval = FALSE}
url_cities <- "https://opendata.arcgis.com/datasets/993d420b9f0742b9afa06622d27a37e0_0.geojson"
path_dest_cities <- file.path("data-raw", "cities")
downloader::download(url = url_cities, destfile = path_dest_cities)
```

```{r cities_sf_raw, eval = FALSE}
cities_raw_sf <-
  path_dest_cities %>%
  sf::st_read() %>% 
  mutate_if(is.factor, funs(as.character)) %>% 
  janitor::clean_names()
```

```{r cities_sf, eval = FALSE}
city_nms <-
  c("Austin", "Corpus Christi", "Dallas", "El Paso", "Houston", "Laredo", "Lubbock", "Odessa", "San Antonio")
cities_sf_filt <-
  cities_raw_sf %>%
  as.data.frame() %>%
  dplyr::select(-geometry) %>%
  filter(city_nm %in% city_nms) %>% 
  arrange(desc(pop2010))

# Reference: https://ryanpeek.github.io/2017-11-05-mapping-with-sf-Part-2/
cities_sf <-
  cities_raw_sf %>% 
  inner_join(cities_sf_filt) %>% 
  mutate(x = map_dbl(geometry, ~st_centroid(.x)[[1]]),
         y = map_dbl(geometry, ~st_centroid(.x)[[2]]))
cities_sf
```

```{r path_cities_sf, eval = FALSE}
path_cities_sf <- 
  file.path(config$dir_data, "cities_sf.shp")
```

```{r cities_sf_export, eval = FALSE}
cities_sf %>% 
  st_write(path_cities_sf, delete_dsn = TRUE)
```

```{r cities_sf_import, eval = FALSE}
cities_sf <-
  st_read(path_cities_sf)
```


```{r hwys_dl, eval = FALSE}
url_hwys <- "https://opendata.arcgis.com/datasets/e661b4004aee4c939569a563b6bb881a_0.geojson"
path_dest_hwys <- file.path("data-raw", "hwys")
downloader::download(url = url_hwys, destfile = path_dest_hwys)
```

```{r hwys_sf_raw, eval = FALSE}
hwys_raw_sf <-
  path_dest_hwys %>%
  sf::st_read() %>%
  mutate_if(is.factor, funs(as.character)) %>% 
  janitor::clean_names()
```

```{r hwys_sf, eval = FALSE}
hwys_sf_n_us <-
  hwys_raw_sf %>%
  as.data.frame() %>%
  dplyr::select(-geometry) %>%
  filter(rte_prfx == "US") %>%
  count(memorial_h, sort = TRUE)

hwys_sf <-
  hwys_raw_sf %>% 
  inner_join(hwys_sf_n_us)
hwys_sf
```

```{r path_hwys_sf, eval = FALSE}
path_hwys_sf <- 
  file.path(config$dir_data, "hwys_sf.shp")
```

```{r hwys_sf_export, eval = FALSE}
hwys_sf %>% 
  st_write(path_hwys_sf, delete_dsn = TRUE)
```

```{r hwys_sf_import, eval = FALSE}
hwys_sf <-
  st_read(path_hwys_sf)
```

```{r schools_tea, eval = FALSE}
schools_tea <-
  teproj::import_ext_csv(
    schools_tea,
    dir = config$dir_data
  )

schools_tea_filt <-
  schools_tea %>%
  filter(year == 2015) %>%
  filter(test == "SAT") %>%
  rename(value = math) %>% 
  # group_by(test, year, county) %>% 
  # summarise_at(vars(value), funs(max(., na.rm = TRUE))) %>% 
  # ungroup() %>% 
  select(test, year, school, district, county, city, value)
```

```{r tx_border, eval = FALSE}
tx_border <- 
  teplot::get_map_data_state(state = "tx") %>% 
  as_tibble()
```

I'll skip over the data collection and munging steps and just show the cleaned data
that I'm using. (See the GitHub repository for the full code.)

```{r sf_show, results = "markup", include = TRUE}
schools_tea_filt %>% glimpse()
schools_sf %>% glimpse()
counties_sf %>% glimpse()
cities_sf %>% glimpse()
hwys_sf %>% glimpse()
tx_border %>% glimpse()
```

Here I create the grid of data that I'll use for the visual. (Thanks
to Jonas for [his example](http://jschoeley.github.io/2018/06/30/bubble-grid-map.html).)

```{r schools_grid, include = TRUE, eval = FALSE}
counties_grid_sf <-
  counties_sf %>%
  st_make_grid(n = c(40, 40))

schools_grid_sf <-
  counties_sf %>%
  left_join(schools_tea_filt) %>%
  select(value) %>%
  # NOTE: Set `extensive = FALSE` to get the mean. Otherwise, set `extensive = TRUE` for the sum.
  st_interpolate_aw(to = counties_grid_sf, extensive = FALSE) %>%
  st_centroid() %>%
  cbind(st_coordinates(.))
```

```{r path_schools_grid_sf, eval = FALSE}
path_schools_grid_sf <- 
  file.path(config$dir_data, "schools_grid_sf.shp")
```

```{r schools_grid_sf_export, eval = FALSE}
schools_grid_sf %>% 
  st_write(path_schools_grid_sf, delete_dsn = TRUE)
```

```{r schools_grid_sf_import, eval = FALSE}
schools_grid_sf <-
  st_read(path_schools_grid_sf)
```

```{r schools_grid_sf_show, include = TRUE, results = "markup"}
schools_grid_sf %>% glimpse()
```

```{r viz_schools_grid, include = TRUE, results = "asis", fig.width = 8, fig.height = 8}
viz_schools_grid <-
  schools_grid_sf %>%
  ggplot() +
  geom_polygon(
    data = tx_border,
    aes(x = long, y = lat, group = group),
    size = 1.5,
    color = "black",
    fill = NA
  ) +
  geom_sf(
    data = hwys_sf,
    linetype = "solid",
    size = 0.1
  ) +
  geom_point(
    aes(x = X, y = Y, size = value, color = value),
    # show.legend = FALSE,
    shape = 16
  ) +
  geom_sf(
    data = cities_sf,
    shape = 16,
    size = 2,
    fill = "black"
  ) +
  ggrepel::geom_label_repel(
    data = cities_sf,
    aes(x = x, y = y, label = city_nm)
  ) +
  coord_sf(datum = NA) +
  scale_color_viridis_c(option = "B", na.value = "#FFFFFF") +
  teplot::theme_map(legend.position = "bottom") +
  labs(
    title = str_wrap("Texas High School Math SAT Scores, 2015", 80),
    caption = "By Tony ElHabr."
  )
viz_schools_grid
```

```{r viz_schools_grid_export, eval = FALSE}
teproj::export_ext_png(
  viz_schools_grid,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 8,
  width = 8
)
```

Cool! I like this visualization because it seems to offer a finer amount
of detail compared to a choropleth.
(In other words, it seems to emphasize specific areas in counties
and not the entire county itself.) Nonetheless, there are some disadvantages
of this technique.

+ There is subjectivity involved in the choice of precision for interpolation.
The grid in my example seems a bit "too" granular around the San Antonio and Austin
area, where it seems like there are no values at all! (Perhaps this
is just an "operator error" on my behalf.)

+ `sf::st_interpolate_aw()` seems to only be capable of aggregating
by sum (with `extensive = TRUE`) or mean (with `extensive = FALSE`). There
are certainly some cases where other aggregation functions would be desirable.
For my example, I actually would have preferred a maximum. An average
is sensitive to area with a relatively small number of schools (that, consequently, may
be "over"-emphasized by the value encoding); and a sum may too strongly
emphasize areas with a large number of schools without providing any insight
into their scores.

For comparison's purposes, let's look at what a choropleth map would look like.
I'll need an additional data.frame for this exercise---`schools_tea_filt_join`---which
is just the `schools_tea_filt` data joined with counties data that
can be retrieved from a call to `ggplot2::map_data()`.

```{r tx_counties, eval = FALSE}
tx_counties <- 
  teplot::get_map_data_county(state = "tx") %>% 
  as_tibble()
```

```{r tx_counties_debug, eval = FALSE}
# NOTE: After trying anti_joins in both directions, noticed that "dewitt" should
# have a space in the `schools_tea` data.
tx_counties %>% 
  anti_join(
    schools_tea_filt %>% 
      mutate_at(vars(county), funs(tolower)) %>% 
      mutate_at(vars(county), funs(if_else(. == "dewitt", "de witt", .))),
    by = c("subregion" = "county")
  ) %>% 
  count(subregion, sort = TRUE)

```

```{r schools_tea_filt_join, eval = FALSE}
schools_tea_filt_join <-
  tx_counties %>% 
  left_join(
    schools_tea_filt %>% 
      group_by(test, year, county) %>% 
      summarise_at(vars(value), funs(mean(., na.rm = TRUE))) %>% 
      ungroup() %>% 
      mutate_at(vars(county), funs(tolower)) %>% 
      mutate_at(vars(county), funs(if_else(. == "dewitt", "de witt", .))),
    by = c("subregion" = "county")
  )
```

```{r schools_tea_filt_join_show, results = "markup"}
schools_tea_filt_join %>% glimpse()
```

```{r viz_schools_chlr, include = TRUE, results = "asis", fig.width = 8, fig.height = 8}
viz_schools_chlr <-
  ggplot() +
  geom_polygon(
    data = tx_border,
    aes(x = long, y = lat, group = group),
    size = 1.5,
    color = "black",
    fill = NA
  ) +
  geom_polygon(
    data = schools_tea_filt_join,
    aes(x = long, y = lat, group = group, fill = value),
  ) +
  geom_sf(
    data = hwys_sf,
    linetype = "solid",
    size = 0.1
  ) +
  geom_sf(
    data = cities_sf,
    shape = 16,
    size = 2,
    fill = "black"
  ) +
  ggrepel::geom_label_repel(
    data = cities_sf,
    aes(x = x, y = y, label = city_nm)
  ) +
  coord_sf(datum = NA) +
  scale_fill_viridis_c(option = "B", na.value = "#FFFFFF") +
  teplot::theme_map(legend.position = "bottom") +
  labs(
    title = str_wrap("Texas High School Math SAT Scores, 2015", 80),
    caption = "By Tony ElHabr."
  )
viz_schools_chlr
```

```{r viz_schools_chlrexport, eval = FALSE}
teproj::export_ext_png(
  viz_schools_chlr,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 8,
  width = 8
)
```

This choropleth actually isn't so bad, but I think I still prefer the bubble grid.

```{r save, include = FALSE}
# NOTE TO SELF: Do not save with blog post or if calling rmarkdown::render().
# save.image(file = path_save)
```

