---
title: An Analysis of Texas High School Academic Competition Results, Part 3 - Individuals
slug: analysis-texas-high-school-academics-3-individuals
date: "2018-05-20"
categories: []
tags:
 - r
 - analysis
 - texas
 - academics
banner: "img/analysis-texas-high-school-academics/viz_comp_stats_byperson_byschool.png"
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  # echo = TRUE,
  echo = FALSE,
  cache = TRUE,
  # include = FALSE,
  fig.align = "center",
  # results = "asis",
  fig.width = 7,
  fig.height = 7,
  # out.width = 7,
  # out.height = 7,
  warning = FALSE,
  message = FALSE
)

```

```{r load, include = FALSE}
library("tidyverse")
library("rlang")
library("teplot")
library("teproj")
library("tetidy")
load(file.path("static", "data", "uil-v02-setup.RData"))
config$path_schools_uil <-
  "https://raw.githubusercontent.com/tonyelhabr/uil-v02/master/data/schools-uil.csv"
schools_uil <-
  config$path_schools_uil %>%
  # teproj::import_path_cleanly()
  readr::read_csv()
config$path_persons_uil <-
  "https://raw.githubusercontent.com/tonyelhabr/uil-v02/master/data/persons-uil.csv"
persons_uil <-
  config$path_persons_uil %>%
  # teproj::import_path_cleanly()
  readr::read_csv()
```

Let's take a look at individual competitors in the academic UIL competitions.

## Individual Participation

The first question that comes to mind is that of participation--which individuals have competed the most?

__NOTE:__

<i>
To give some context to the values for individual participants, 
I'll include the numbers for myself ("Elhabr, Anthony") in applicable contexts.
</i>

```{r person_n_top, include = FALSE, dependson = c("load")}
persons_n <-
  persons_uil %>%
  group_by(name, school, city, conf) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  tetidy::rank_arrange_at("n")
persons_n
```

```{r html_person_n, echo = FALSE, include = TRUE, dependson = c("load", "persons_n")}
html_persons_n <-
  persons_n %>%
  create_kable_filt_at(
    config = config,
    col_rgx = "name"
  )
html_persons_n
```

Although the names here may not provide much insight, the counts provide some context
regarding the limits of individual participation.

Given that counts of overall participation may not be indicative of anything directly,
it may be a better idea to break it down by conference.

```{r persons_n_top_byconf, include = FALSE, dependson = c("persons_n")}
# n_top_conf <- 10000
# n_top_conf <- nrow(persons_n)
persons_n_top_byconf <-
  persons_n %>%
  # slice(1:n_top_conf) %>%
  group_by(conf) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  tetidy::rank_arrange_at("n")
persons_n_top_byconf
```

```{r html_person_n_top_byconf, echo = FALSE, include = TRUE, dependson = c("load", "persons_n_top_byconf"), eval = FALSE}
html_persons_n_top_byconf <-
  persons_n_top_byconf %>%
  mutate_at(vars(n), funs(scales::comma)) %>%
  teproj::create_kable()
html_persons_n_top_byconf
```

```{r viz_persons_n_top_byconf, echo = FALSE, include = TRUE, fig.height = 5, fig.width = 7, dependson = c("load", "persons_n")}
viz_persons_n_top_byconf <-
  persons_n %>%
  ggplot(aes(x = conf, y = n)) +
  geom_col() +
  teplot::scale_y_pretty_comma() +
  teplot::theme_te_dx()
  labs(
    title = "Count of Individual Scores Recorded for Each Conference",
    caption = config$viz_footer,
    x = "Conference",
    y = "Count of Individual Scores"
  )
viz_persons_n_top_byconf
teproj::export_ext_png(
  viz_persons_n_top_byconf,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 5,
  width = 7
)
```

It seems that there has not been as much invdividual participation in the 6A conference (`conf 6`)--
which is the conference with largest high schools (according to student body size).

I hypothesize that this phenomenon
can be attributed to "pre-filtering of talent" by these large schools.
In other words, conference 6A schools may be more likely to designate their individual
competitors to compete in only specific competitions and prevent any student
who may be capable, yet not fully prepared, from entering a competition.
High standards and expectations of aptitude are relatively common at very large schools,
even if what may be deemed "unacceptable" at such a school would be very satisfactory
at a smaller school.
By comparison, schools in all other conferences may be more willing to let individual
students compete in as many competition types as they desire, even if they have
not prepared for them whatsoever.

Such a phenomenon might be evident in lower scores (in aggregate) for conferences where participation
is greater. In fact, this is exactly what is observed. [^previous] On average, 
conference 6A has the highest scores, while conference 1A has the lowest.

[^previous]:
See the previous post.

```{r viz_persons_stats_byconf, include = FALSE, fig.height = 7, fig.width = 7, eval = FALSE}
# TODO: Figure out a good way to include the previously generated visual.
viz_persons_stats_bycompconf
```

So what about people's scores? Who did best according to score?
In order to simplify the data, let's look at a couple of statistics
based on score, aggregating across all scores for each individual.
In particular, let's look at the average and sum of placing percent rank (`prnk`) and of
individual competitors "defeated" (`n_defeat`). (Note that
competitors defeated is defined as 
the number of scores less that that of a given individual for a unique competition,
and a unique competition is defined as a unique combination of 
year, competition level, and competition type.)

```{r stats, include = FALSE, dependson = c("load")}
persons_stats <-
  persons_uil %>%
  add_stats_cols_by_at(cols_grp = c("name", "school", "city"))

schools_stats <-
  schools_uil %>%
  add_stats_cols_by_at(cols_grp = c("school", "city"))
```


```{r html_persons_stats, echo = FALSE, include = TRUE, dependson = c("load", "stats")}
html_persons_stats <-
  persons_stats %>%
  select(-n_advanced_sum, -n_state_sum) %>%
  select(-prnk_sum, -n_defeat_sum) %>% 
  mutate_if(is.numeric, funs(round(., 2))) %>% 
  create_kable_filt_at(
    config = config,
    col_rgx = "name"
  )
html_persons_stats
```

Also, I think it's interesting to look at the distribution of counts for
competitors defeated, advancement, and state competition appearances. The heavily right skewed
distribution of values gives an indication of the difficulty of succeeding consistently.

```{r viz_persons_stats_n_x, echo = FALSE, include = TRUE, fig.height = 7, fig.width = 7, dependson = c("load", "html_persons_stats")}
count_rename_at <-
  function(data = NULL, col = NULL, col_out = "value") {
    col <- sym(col)
    col_out <- sym(col_out)
    data %>%
      count(!!col) %>% 
      rename(!!col_out := !!col)
  }
do_count_rename_at <-
  function(data = NULL) {
    bind_rows(
      # data %>%
      #   count_rename_at("n_defeat_sum") %>% 
      #   mutate(lab_stat = "Competitors Defeated"),
      data %>%
        count_rename_at("n_advanced_sum") %>% 
        mutate(lab_stat = "Competition Advancements"),
      data %>%
        count_rename_at("n_state_sum") %>% 
        mutate(lab_stat = "State Competitions")
    )
  }
visualize_stats_n_x <-
  function(data = NULL) {
    data %>% 
      ggplot(aes(x = value, y = nn)) +
      geom_col() +
      facet_wrap(~lab_stat, scales = "free") +
      scale_y_log10(labels = scales::comma) +
      # teplot::scale_xy_pretty_comma() +
      teplot::theme_te_facet_dx() +
      theme(legend.position = "none") +
      labs(
        x = "Count of Occurrences",
        y = "Log 10 of Count"
      )
  }

stats_n_x <-
  bind_rows(
    persons_stats %>%
      do_count_rename_at() %>% 
      mutate(entity = "persons"),
    schools_stats %>%
      do_count_rename_at() %>% 
      mutate(entity = "schools")
  )

lab_title_persons_stats_n_x <-
  "Distribution of Counts of Individual Performance Metrics"
viz_persons_stats_n_x <-
  stats_n_x %>% 
  filter(entity == "persons") %>% 
  visualize_stats_n_x() +
  labs(title = lab_title_persons_stats_n_x, caption = config$viz_footer)
viz_persons_stats_n_x
teproj::export_ext_png(
  viz_persons_stats_n_x,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 7,
  width = 7
)
```

For comparison's sake, let's visualize the same metrics aggregated at the school level.
Keep in mind that while the sample of students
should have larger counts for number of advancements and state competition appearances (y-axis) 
for any given number of occurences (x-axis) because there are many more students than schools,
schools are more likely to have a wider range of occurrences (x-axis) 
because there are less
schools in each competition (compared to the number of individuals).

To understand why this is true, let's take an example:
Say there is a District level competition where there are 8 
schools and 40 individuals competing. It is more likely that a given school advances to the 
next level of competition (as a result of having a total score that is higher than the scores
of the other 7 schools) than
any single individual, who if not from the school that advances, can only
advance as a result of having a top "n" (e.g. 3) score.

```{r viz_schools_stats_n_x, echo = FALSE, include = TRUE, fig.height = 7, fig.width = 7, dependson = c("load", "viz_persons_stats_n_x")}
viz_schools_stats_n_x <-
  stats_n_x %>% 
  filter(entity == "schools") %>% 
  visualize_stats_n_x() +
  labs(title = str_replace(lab_title_persons_stats_n_x, "Individual", "School"), caption = config$viz_footer)
viz_schools_stats_n_x
teproj::export_ext_png(
  viz_schools_stats_n_x,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 7,
  width = 7
)
```

We see that the distributions are skewed towards the right here as well, although
not quite as "evenly". This indicates that some schools tend to perform well
at a more consistent rate than individuals themselves. Intuitively, this makes sense.
It can be very difficult for individuals alone to beat out the competition, especially
if they have an "off" day.
On the other hand, schools, relying on teams of individuals, are placed according to the sum
of the top "n" (e.g. 3) of individual competitor scores.
Thus, because school scores are dependent on groups of individuals--
who will tend to perform more consistently in aggregate than any one individual--
school placings are more likely to be similar across years, meaning that schools
that are observed to do well in any given year are more likely to do well in other years as well
(relative to individual competitors).

So it should be obvious that it is difficult to make it the highest level of competition--State.
But exactly how difficult is it? Let's identify those people (and their scores)
who have made the State competition level four times--which
is the upper limit for a typical high school student [^eligibility]--
for a given competition type.

[^eligibility]:
The assumption here is that each student takes four years to complete high school.

```{r persons_multi, include = FALSE, dependson = c("load")}
persons_multi <-
  persons_uil %>%
  count(name, school, city, conf, comp, complvl, complvl_num) %>% 
  filter(n > 1)
```

```{r persons_n4, include = FALSE, dependson = c("load", "person_multi")}
persons_n4 <-
  persons_multi %>% 
  filter(n == 4)
```

```{r html_persons_n4_state, echo = FALSE, include = FALSE, dependson = c("load", "persons_n4")}
html_persons_n4_state <-
  persons_n4 %>%
  filter(complvl == "State") %>% 
  distinct(name, school, city, conf, comp) %>% 
  create_kable_filt_at(
    config = config,
    col_rgx = "name"
  )
html_persons_n4_state
```

Clearly, these individuals represent a very small subset of the total sample. 
They might be considered the "elite". Of these individuals, who has appeared
in State competitions for more than one type of competition?

```{r html_persons_n4_state_n2}
html_persons_n4_state_n2 <-
  persons_n4 %>%
  filter(complvl == "State") %>% 
  distinct(name, school, city, conf, comp) %>% 
  count(name, school, city, conf, sort = TRUE) %>% 
  filter(n > 1) %>% 
  create_kable_filt_at(
    config = config,
    col_rgx = "name",
  )
html_persons_n4_state_n2
```

I would consider those individuals appearing here to be the "elite of the elite".

## Individual Performance

Now, I want to try to answer a somewhat ambiguous question:
Which individuals were most "dominant"?

### Evaluating "Dominance"

Because the term "dominance" is fairly subjective, it must be defined explicitly.
Here is my definition/methodology, along with some explanation.

First, I assign a percent rank to individual placings in all
competitions based on score relative to other scores in that competition. 
I choose to use percent rank--which is a always a value between 0 and 1--because
it inherently accounts for the
wide range of number of competitors across all competitions. 
(For this context, a percent rank of 1 corresponds to the highest score in a given
competition [^competition], and, conversely,
a value of 0 corresponds to the lowest score.)

[^competition]:
A unique competition is defined as one having a unique year, competition type, and competition level.

I should note that I evaluated some other metrics for gauging individual
success, including the total number
of individuals defeated in competitions. 
Percent rank based on score and number of defeats attempt to quantify the same underlying variable,
but I think percent rank is a little more "natural" to interpret because
it contextualizes number of competitors with its unit range.
By comparison, the interpretation of number of defeats is less direct because
the number of other competitors is not accounted directly.

Then, to come up with a final set of ranks, one for each unique competitor,
based on the percent ranks for individual competitions,
I simply sum up the percent ranks for each individual.

The sum is used instead of an average [^average] because rankings based on averages
--and inferences made upon them--are sensitive
to individuals who do not compete in many competitions, yet place very well in them.
A final ranking based on a summed value does not suffer from this pitfall,
although it can be sensitive to the sample size of each participant. (i.e. 
An individual might participate in a high number of competitions and
under-perform relative to the average in all of them, yet their final ranking, 
based on summed percent ranks, might indicate that they are an above-average performer.)

[^average]:
The average may also be considered a valid means of aggregating the values for each individual.


```{r persons_rnks, include = FALSE, dependson = c("load")}
add_rnks_cols_by_at <-
  function(data = NULL,
           col = NULL,
           cols_grp = c("name", "school", "conf")) {
    stopifnot(is.character(col), length(col) == 1, length(intersect(names(data), col)) == 1)
    stopifnot(is.character(cols_grp),
              length(intersect(names(data), cols_grp)) == length(cols_grp))
    data %>%
      group_by(!!!syms(cols_grp)) %>%
      summarise_at(vars(!!sym(col)), funs(sum, mean)) %>%
      ungroup() %>%
      mutate(rnk_sum = row_number(desc(sum)),
             rnk_mean = row_number(desc(mean))) %>% 
      arrange(rnk_sum, rnk_mean)
  }

persons_rnks <-
  left_join(
    persons_uil %>%
      add_rnks_cols_by_at("prnk"),
    persons_uil %>%
      add_rnks_cols_by_at("n_defeat") ,
    by = c("name", "school", "conf"),
    suffix = c("_prnk", "_n_defeat")
  ) %>% 
  select(matches("name|school|conf|^rnk_"))
persons_rnks
```

```{r html_person_rnks, echo = FALSE, include = TRUE, dependson = c("load", "persons_rnks")}
html_persons_rnks <-
  persons_rnks %>%
  create_kable_filt_at(
    config = config,
    col_rgx = "name"
  )
html_persons_rnks
```

Some of the same individuals from the participation-based ranking
of competitors also appear among the top of the ranks by my evaluation of domination.
This is somewhat expected due to the nature of my methodology (in particular, my
choice to use a sum instead of an average or some other statistic).
Certainly some additional statistical analysis could be done here to investigate
other methods of quantifying dominance (beyond my analysis). [^more]
Because there is no well-agreed upon metric or method anywhere for quantifying
dominance for this particular topic, it's difficult to really judge the findings here.

[^more]:
I'll leave further analysis for another person and/or time.

In any matter, the difference between the ranks based on the average and the sum
of percent ranks is not all too great--I found that the correlation between the
two is anywhere between ~0.75 and ~0.95 when using either score percent rank
or number of defeats as the metric for ranking.

```{r persons_rnks_cors, include = FALSE, dependson = c("load", "persons_rnks")}
correlate_rnks_at <-
  function(data = NULL, cols = str_subset(names(data), "^rnk_")) {
    data %>%
      select(one_of(c(cols))) %>%
      corrr::correlate()
  }
persons_rnks_cors <-
  persons_rnks %>% 
  select(starts_with("rnk")) %>%
  corrr::correlate()
persons_rnks_cors
```

```{r html_persons_rnks_cors, echo = FALSE, include = TRUE, dependson = c("load", "persons_rnks_cors")}
html_persons_rnks_cors <-
  persons_rnks_cors %>%
  mutate_if(is.numeric, funs(round(., 4))) %>% 
  teproj::create_kable()
html_persons_rnks_cors
```

### Evaluating "Carrying"

Another way of identifying superb ability is to compare the scores of individuals
with those of their teammates. Individuals with high scores relative to 
their teammates might be said to have "carried" their teammates.
Although this kind of evaluation is dependent on the skill of each team
(independent of the competition setting), I think that it is another interesting way of
evaluating skill.

```{r persons_stats_byschool, include = FALSE, dependson = c("load")}
n_ingroup_min <- 3
persons_stats_byschool <-
  persons_uil %>%
  # filter(school == config$rgx_school_filt, year <= 2012, year >= 2011) %>%
  group_by(comp, complvl, conf, year, school) %>%
  mutate(n_ingroup = n()) %>%
  filter(n_ingroup >= n_ingroup_min) %>%
  mutate(ingroup_rnk = row_number(desc(score))) %>%
  mutate(ingroup_score_other = (sum(score) - score)  / (n_ingroup - 1)) %>%
  mutate(ingroup_diff = score - ingroup_score_other) %>%
  ungroup() %>%
  mutate(among_group_rnk = row_number(desc(ingroup_diff))) %>%
  arrange(among_group_rnk)
persons_stats_byschool
```

```{r viz_comp_stats_byperson_byschool, echo = FALSE, include = TRUE, fig.height = 7, fig.width = 7, dependson = c("load", "persons_stats_byschool")}
set.seed(42)
pct_top_ingroup_diff_prnk <- 0.05
lab_caption_comp_stats_byperson_byschool <-
  paste0(
    config$viz_footer,
    str_wrap(
      paste0(
        "Blue points highlight my personal scores. ",
        sprintf("Red emphasizes top %.0f%% of individuals who \"carried\" ", 
                100 * pct_top_ingroup_diff_prnk),
        "their teammates, according to differences in percent rank ",
        "of difference in individual score and average of teammates scores. ",
        "Only unique school-competitions observations having ",
        sprintf("at least %.0f team members are considered.", n_ingroup_min)
      ), 100
    )
  )

viz_comp_stats_byperson_byschool <-
  persons_stats_byschool %>%
  mutate(ingroup_diff_prnk = percent_rank(ingroup_diff)) %>%
  mutate(ingroup_diff_toppct = if_else(ingroup_diff_prnk >= (1 - pct_top_ingroup_diff_prnk), TRUE, FALSE)) %>%
  sample_frac(size = 0.05) %>%
  # ggplot(aes(x = score, y = group_score)) +
  ggplot(aes(x = score, y = ingroup_score_other)) +
  geom_point(aes(color = ingroup_diff_toppct), alpha = 0.1) +
  geom_point(
    data =
      persons_stats_byschool %>%
      filter(str_detect(name, config$rgx_name_filt)),
    aes(x = score, y = ingroup_score_other),
    color = "blue",
    size = 5,
    shape = 18
  ) +
  scale_color_manual(values = c("black", "red")) +
  geom_abline(
    aes(intercept = 0, slope = 1),
    color = "black",
    linetype = "solid",
    size = 2
  ) +
  geom_abline(
    aes(
      intercept = -quantile(ingroup_diff, 1 - pct_top_ingroup_diff_prnk, na.rm = TRUE),
      slope = 1
    ),
    color = "red",
    size = 1,
    linetype = "dashed"
  ) +
  labs(
    x = "Individual Score",
    y = "Average Score of Teammates",
    title = "Distribution of Individual and Teammate Scores",
    caption = lab_caption_comp_stats_byperson_byschool
  ) +
  teplot::theme_te_facet() +
  theme(legend.position = "none")
viz_comp_stats_byperson_byschool

teproj::export_ext_png(
  viz_comp_stats_byperson_byschool,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 7,
  width = 7
)
```

I was hoping that I might see myself appearing among the most dominant by this measure
of skill, but it does not say anything necessarily bad about myself that I don't.
I competed with other individuals who I considered to be very knowledgeable and who
often scored better than me. Also, from the opposite point of view, I don't think
I was a poor performer who relied upon teammates to boost the team's overall score.
This is why the data points corresponding to me show up in the middle of the pack
in the previous visual.

### "Improvement"

I think something else that would be interesting to look at is personal "improvement"
between years. Theoretically, if we assume that individuals improve their academic
ability and competition skills every year, then we should see individual scores
for a given competition type and level increase from one year to next.
I would be very surprised to find that this is not true.

To evaluate improvement, we can simply reduce the whole data set to just those
who have competed in the same competition type and level in more than one year
and check whether or not their scores increased or decreased from one year to the next. [^years]
Actually, in order to account for variance in competition difficulty across years,
it's better to use the percent rank of the individual's placing (based on score) rather than
score itself.

[^years]:
I don't think it is relevant to require that the scores be in consecutive years,
so I don't enforce that criteria.

```{r persons_improve_n, include = FALSE, dependson = c("load", "persons_multi")}
persons_improve <-
  persons_uil %>% 
  semi_join(persons_multi) %>% 
  group_by(name, school, city, conf, comp, complvl, complvl_num) %>% 
  arrange(year, .by_group = TRUE) %>% 
  # mutate(order_byperson = row_number(desc(year))) %>% 
  # mutate(prnk_lag = lag(prnk)) %>% 
  mutate(improve = ifelse(prnk > lag(prnk), TRUE, FALSE)) %>% 
  ungroup() %>% 
  filter(!is.na(improve))

persons_improve_n <-
  persons_improve %>% 
  count(improve)
```

```{r html_persons_improve_n, echo = FALSE, include = TRUE, dependson = c("load", "persons_improve_n")}
html_persons_improve_n <-
  persons_improve_n %>% 
  mutate_if(is.numeric, scales::comma) %>% 
  create_kable_filt_at(
    config = config,
    col_rgx = "name"
  )
html_persons_improve_n
```

So it is true that individual scores--actually, percent rank of placings--do tend
to improve as the individual ages. But is this trend statistically significant?
That's easy enough to answer--we can simply perform a 
[binomial test](https://en.wikipedia.org/wiki/Binomial_test)--where the 
[null hypothesis](https://en.wikipedia.org/wiki/Null_hypothesis)
is that the distribution of "TRUE" and "FALSE" regarding improvement
is truly a 50 % - 50 % split. If the 
[p-value](https://en.wikipedia.org/wiki/P-value) of the test is below a threshold value--let's
say 0.05--then we can deny the null hypothesis and say that there is a non-trivial
trend of individual improvement.

```{r persons_improve_btest, include = FALSE, dependson = c(-1)}
persons_improve_n_c <- tibble::deframe(persons_improve_n)
persons_improve_btest <-
  binom.test(persons_improve_n_c["TRUE"], sum(persons_improve_n_c)) %>% 
  broom::tidy()
persons_improve_btest
```

```{r html_persons_improve_btest, echo = FALSE, include = TRUE, dependson = c(-1)}
html_persons_improve_btest <-
  persons_improve_btest %>%
  select(estimate, p.value, conf.low, conf.high) %>% 
  mutate_if(is.numeric, funs(round(., 4))) %>% 
  gather(metric, value) %>% 
  teproj::create_kable()
html_persons_improve_btest
```

In fact, this is exactly what is observed.

Now, let's reduce the set to just those who have appeared in competitions at a given
competition type four times (for the sake of visualization)
and plot the scores across years for each individual.

```{r viz_persons_n4, echo = FALSE, include = TRUE, fig.height = 7, fig.width = 7, dependson = c("load", "persons_n4")}
persons_n4_viz <-
  persons_uil %>% 
  semi_join(persons_n4) %>% 
  # filter(complvl == "State") %>% 
  unite(id, name, school, city, conf, comp, complvl, complvl_num, sep = "-", remove = FALSE) %>% 
  group_by(id, complvl) %>% 
  mutate(year_order = row_number(year)) %>% 
  ungroup()

lab_title_person_n4 <-
  str_wrap("Percent Rank of Scores for Individuals Competing Four Times in a Specific Competition", 60)
lab_caption_persons_n4 <-
  paste0(
    config$viz_footer,
    str_wrap(
      paste0(
        "Grey lines: individual scores across sequence of years.",
        "Blue lines: linear regression model. Blue points: average across all individuals."
      )
    ), 100
  )

viz_persons_n4 <-
  persons_n4_viz %>% 
  ggplot(aes(x = year_order, y = prnk)) +
  geom_line(aes(group = id), alpha = 0.1) +
  geom_smooth(method = "lm", color = "blue") +
  # geom_smooth(method = "loess", color = "red") +
  geom_point(
    data = 
      persons_n4_viz %>%
      group_by(complvl, year_order) %>%
      summarise(prnk_mean = mean(prnk)),
    aes(y = prnk_mean),
    size = 3,
    color = "blue"
  ) +
  facet_wrap(~complvl, scales = "free") +
  teplot::theme_te_facet() +
  labs(
    title = lab_title_person_n4,
    caption = lab_caption_persons_n4,
    x = "Order of Year",
    y = "Percent Rank of Score"
  )
viz_persons_n4
teproj::export_ext_png(
  viz_persons_n4,
  export = config$export_viz,
  dir = config$dir_viz,
  units = "in",
  height = 7,
  width = 7
)

```

It is evident visually that people do have a tendency to improve over time.

## Wrap-up

I'll leave the discussion of individuals at that, although there is much
more that could be explored.


