---
title: Investigating Ranks, Monotonicity, and Spearman's Rho with R
slug: ranks-monotonicity-spearman-rho-r
date: "2018-04-16"
categories: []
tags:
 - r
 - rank
 - monotonicity
 - spearman
 - dplyr
 - ggplot2
 - broom
banner: "img/ranks-monotonicity-spearman-rho-r/banner.png"
---

```{r setup, echo = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  echo = TRUE,
  # echo = FALSE,
  cache = TRUE,
  include = FALSE,
  # results = "markdown",
  results = "hide",
  fig.align = "center",
  # fig.show = "asis",
  fig.width = 6,
  fig.height = 6,
  # out.width = 6,
  # out.height = 6,
  warning = FALSE,
  message = FALSE
)

```

```{r params}
params <-
  list(
    dir_data = file.path("static", "data"),
    dir_viz = "figs",
    export_viz = FALSE
  )
```


```{r packages}
library("dplyr")
library("ggplot2")
library("tidyr")
library("teplot")
```

```{r setup_vars_funcs}
# library("ggplot2")
theme_base <- function() {
  teplot::theme_te_b_facet(
    base_size = 12,
    plot_title_size = 18,
    subtitle_size = 16,
    axis_title_size = 12,
    caption_face = "plain",
    caption_size = 12,
    plot_margin = margin(2, 2, 2, 2)
  ) +
  theme(
    legend.text = element_text(size = 12)
  )
}

theme_tile <-
  function() {
  theme_base() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 90)
    )
}
labs_tile <-
  function() {
    labs(x = NULL, y = NULL)
  }

create_kable <-
  function(data = NULL, num_show = 10) {
  
  num_rows <- nrow(data)
  show_fn <- ifelse(num_rows > num_show, TRUE, FALSE)
  if(show_fn) {
    data <- data %>% slice(1:num_show)
  }
  
  out <-
    data %>% 
    knitr::kable( "html", escape = FALSE) %>%
    kableExtra::kable_styling(full_width = FALSE, position = "center")
  
  if(show_fn) {
    out <-
      out %>%
      kableExtra::add_footnote(c(sprintf("# of total rows: %.0f", num_rows)), notation = "number")
  }
  out
}
```

## The Problem

I have a bunch of data that can be categorized into many small __groups__.
Each small group has a set of __values__ for an ordered set of __intervals__.
Having observed that the values for most groups seem to increase
with the order of the interval,
I hypothesize that their is a statistically-significant,
monotonically increasing trend.

### An Analogy

To make this abstract problem more relatable, imagine the following scenario.

There are many companies (__groups__) selling products in 
an industry that is on the verge of widespread growth.
Each company sets a projection (not a goal) for end-of-year sales (__value__).
and adjust this projection once per quarter (i.e. four times a year) (__interval__)
after analyzing their to-date sales.
Having observed that
most companies that follow this scheme tend to increase
their goals for next-quarter sales (based on an initial, beginning-of-the-year
projection that is too low),
the market analyst (me) wonders if there is a non-trivial, positive trend across quarters in the year. [^hypothesis]:


[^hypothesis]:
The market analyst does not necessarily hypothesize
why annual projections tend to be high (i.e. perhaps due to over-confidence)

To summarize how the "variables" in this analogy relate to my abstract description of the situation,

+ groups = companies
+ value = sales
+ interval = quarter-year

(Actually, there is one more variable in this context, given that I am interested in
relative value---the __rank__ of the value, relative to the other values.)

And, to make the example concrete, here is what the data (for four companies) might look like.

```{r data_ex}
set.seed(42)
n <- 4L
ya <- sort(round(runif(n, min = 1L, max = 10L), 1), decreasing = FALSE)
yb <- sort(round(runif(n, min = -1L, max = 1L), 2), decreasing = FALSE)
yc <- sort(sample(100L:10000L, size = n, replace = FALSE), decreasing = FALSE)
yd <- sort(sample(-100L:100L, size = n, replace = FALSE), decreasing = FALSE)

# Purposely make some groups non-monotonic.
yc <- yc[c(2, 1, 2, 4)]
yd[3] <- 400

data_ex <-
  tibble::tibble(
    company = sort(rep(LETTERS[1:n], n)),
    quarter = rep(1L:n, n),
    sales = c(ya, yb, yc, yd)
  )

```

```{r data_ex_show, include = TRUE, echo = FALSE, results = "markup"}
data_ex %>% 
  mutate_at(vars(quarter), funs(paste0("q", .))) %>% 
  tidyr::spread(quarter, sales) %>% 
  mutate_if(is.numeric, funs(as.character)) %>% 
  create_kable()
```

```{r viz_ex, include = TRUE, echo = FALSE, fig.width = 5, fig.height = 5}
data_ex %>% 
  ggplot(aes(x = quarter, y = sales, color = company)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm") +
  scale_y_continuous(labels = scales::dollar) +
  teplot::scale_color_te() +
  facet_wrap(~company, scales = "free_y") +
  theme_base() +
  theme(legend.position = "none", panel.background = element_rect()) +
  labs(
    title = "Sales Per Quarter",
    caption = "Note that the \"free\" y-axis helps overcome the difference in magnitude among values."
  )
```


### (Bad) Solution: Linear Regression

Just at first thought, running a univariate
[linear regression](https://en.wikipedia.org/wiki/Linear_regression) might seem like a good way
of attacking this problem.
However, there are a couple of basic "gotchas" that make ordinary linear regression
a not-so-great idea for this scenario:

+ There are not many intervals (i.e. independent variables) per group.
(This condition inherently makes any kind of model---not just a linear regression one---sensitive
to the samples. In general, there is a problem with creating models with
[small sample sizes](https://garstats.wordpress.com/2017/02/04/small-sample-sizes/).)

+ The values across groups might have very different magnitudes.
(Thus, trying to create a single, general model that groups all of the data and
uses the group labels as a categorical independent variable would likely lead to unreliable results.)

+ The values themselves might be very volatile for a single group.
(This might be interpreted as a violation of the
[normal distribution assumption of linear regression](https://www.statisticssolutions.com/assumptions-of-linear-regression/).
Additionally, if the values are truly monotonic, the assumption of no autocorrelation
might also be violated.)

Aside from these caveats,
the value for a given interval is not relevant---rather,
its relationship with all other values is, and, more specifically,
its relationships with the previous and subsequent values. [^prediction]

[^prediction]:
Additionally, prediction is not the concern---rather, quantification of trend is.
(While regression certainly can help with trend identification,
its capability to create predictions is perhaps its better use.)

### (Better) Solution: _Spearman's Rho_

Given the nature of the data (which one might say is non linear) and my
intent to quantify ordinality between two variables, it turns out that
[Spearman's rho](https://en.wikipedia.org/wiki/Spearman%27s_rank_correlation_coefficient),
in theory, provides exactly the measurement that I want---it quantifies the association
between paired samples using the ranks of the variables (not their values) relative to their samples.
Notably, the statistical significance (i.e. via a p-value) can be calculated
(traditionally, using a [t-test](https://en.wikipedia.org/wiki/Student%27s_t-test)),
which should be handy for my intent on identifying non-triviality.

```{r n_iris}
n_iris <- 10
```

Nonetheless, even though this metric seems promisable, it will certainly be
sensitive to the small samples of each group (assuming that it is calculated
for each group).
Don't believe me? Check out how the Spearman's rho value changes
for the numerica columns in the built-in `iris` dataset (which has `r nrow(iris)` rows) when
it is calculated for just the first `r n_iris` rows.

```{r iris_corrr, include = TRUE, echo = FALSE, fig.height = 6, fig.width = 8}

show_iris_ggcorr <- function(data = NULL, lab_subtitle = NULL) {
  
  n <- nrow(data)
  data %>% 
    select_if(is.numeric) %>% 
    GGally::ggcorr(
      method = c("pairwise", "spearman"),
      # geom = "circle",
      legend.position = "none",
      low = "cyan",
      mid = "grey95",
      high = "red",
      label = TRUE,
      label_size = 5
    ) +
    theme(plot.subtitle = element_text(size = 16)) +
    # labs(subtitle = paste0(n, " Rows"))
    labs(subtitle = lab_subtitle)
}

grobs_iris_ggcorr <-
  bind_rows(
    iris %>% mutate(grp = paste0("All ", nrow(iris), " Rows")),
    iris %>% mutate(grp = "First 10 Rows") %>% slice(1:n_iris)
  ) %>% 
  group_by(grp) %>%
    select_if(is.numeric) %>%
  nest() %>%
  mutate(gg = purrr::map2(data, grp, show_iris_ggcorr))
grobs_iris_ggcorr <-
  gridExtra::arrangeGrob(
    grobs = grobs_iris_ggcorr$gg,
    ncol = 2,
    top = grid::textGrob("Spearman's Rho for `iris` Dataset",
                   gp = grid::gpar(fontsize = 18,
                             family = "text",
                             font = 2)
    )
  )
grid::grid.newpage()
viz_iris_ggcorr <- grid::grid.draw(grobs_iris_ggcorr)

```

```{r iris_corrr_2, eval = FALSE}
corestij <- function(i, j, data) {
  cor.test(data[, i], data[, j], method = "spearman")$estimate
}
corpij <- function(i, j, data) {
  cor.test(data[, i], data[, j], method = "spearman")$p.value
}
corest <- Vectorize(corestij, vectorize.args = list("i", "j"))
corp <- Vectorize(corpij, vectorize.args = list("i", "j"))
n <- ncol(iriis)
cors <-
  outer(1:n, 1:n, corest, data = iriis) %>% 
  tibble::as_tibble() %>% 
  purrr::set_names(names(iriis)) %>% 
  mutate(rowname = names(iriis)) %>% 
  select(rowname, everything()) %>% 
  gather(col, value, -rowname)
cors %>% 
  ggplot(aes(x = rowname, y = col)) +
  geom_tile(aes(fill = value)) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "grey95", midpoint = 0) +
  coord_equal() +
  theme_base() +
  theme_tile() +
  labs_tile()
```


### Another Solution: Custom Heuristic

So, what can be done? Well,
even with the hope that the Spearman's rho metric provides for quantification and significance inference,
I thought that I would try to create some kind of easily understandable heuristic
that I could explain to someone else 
without having to delve into statistical theory.
Nonetheless, I would be ignorant to not compare (and validate) the results of my heurisitc
with those of statistical theory after creating my heurisitc.

Having this plan in mind, I began to think about how I would define my heuristic,
which, in essence, tries to quantify [__monotocity__](https://en.wikipedia.org/wiki/Monotonic_function).
But what exactly constitutes monoticity? Surprisingly, that's a more complex
question than it might seem. [^mntc_complexity]
(For example, did you know that numbers may be strictly or weakly monotonic?)

[^mntc_complexity]:
I'll leave the reader to dive into all of the theory.

For my purposes, I don't necessarily care if the set of values
is _strictly_ increasing or decreasing, but they should be 
"sufficiently" increasing or decreasing.
For example, while it is clear that the sequence `1`, `2`, `3`, `4` is strictly monotonic (increasing)
and the sequence `1`, `2`, `4`, `3` is not, I would consider the latter "sufficiently" monotonic.
On the the hand, 
I would consider something like  `4`, `2`, `3`, `1` because the `1` and `4` are
"badly" misplaced in one another's appropriate places, which are at the extreme ends of the sequence.
Moreover, if I was intent on identifying
increasing monoticity (as opposed to decreasing monotonicity),
I would consider `4`, `3`, `2`, `1` "bad", even though it is strictly
monotonically decreasing. But what about something like `1`, `4`, `3`, `2` (again, assuming
that I am evaluating increasing monotonicity)? Even though the
`2` and `4` are swapped, I might still consider this sequence "sufficiently" monotonic 
because the `1` and `3` are placed correctly and the `2` and `4` are "not too far apart".
Anyways, it's easy to see how having some kind of formal definition/calculation/criteria for
monotonicity is handy.

#### The "Algorithm"

After some thinking,
I came up with the following algorithm (if one can even call it that).

(__NOTE:__ I explicitly list the variable names that I use in the code that follows to help
the reader understand the relationships between these steps and the implementation.)

1. Given an `n`-length sequence of arbitrary values, assign each value
an integer value between `1` and `n` to represent its "actual" rank. This rank
is to be assigned based on relative value in
the set of [real numbers](https://en.wikipedia.org/wiki/Real_number). [^rank_init]

+ In the machine learning setting, this set of ranks is the 
dependent variable (i.e. `y`) to predict.
+ In the example situation described before, it equates to the rank that
would be assigned to the quarterly interval based on sales relative to the other quarters.
+ In the simulation that I run below, this is the variable `y0`.

2. Create a matrix of all permutations of actual rank and "assigned" rank.
To be clear, this "assigned" rank is independent of the actual rank and value.
To make things straightforward, these assigned ranks should be transformed to
use only the same set of possible rank values dictated by the actual ranks (i.e. integers
between 1 and `n`).

+ In the machine learning setting, this "assigned" rank is the 
independent variable (i.e. `x`) used as a predictor.
+ In the example, it is the quarterly interval.
+ In the simulation that follows, assigned rank is `x0`, and the matrix (actually, a `tibble`)
of combinations is `data_permn`.

3. Calculate the absolute difference between the "actual" and "assigned" ranks
for each value in the sequence.
Subtract this distance from the maximum rank value in the sequence. The resulting
value is what I call the
"inverse monotonic distance" (`mntc_distinv` in the following code).

4. Repeat the calculation of inverse monotonic distance for all groups (`grp`) of
"actual" (`y0`) and "assigned" (`x0`) ranks.

5. Sum up the inverse monotonic distance for each value in the permutation group and
take the average of this sum for each group. [^sum_or_avg]
Re-scale this per-group value to a 0 to 1 basis. [^re-scale] (In the code that follows,
I re-use the variable name `mntc_distinv` for this transformed value.)

6. Identify any group (`grp`) corresponding to a sum-averaged-re-scaled value
(`mntc_distinv`) in the upper 50% quantile of all values
(i.e. assigned the value `"(0.5,1]"` for the `mntc_tier2` variable)
as __"sufficiently" monotonic__. (The `mntc_tier2` variable can be interpreted as my heuristic.)

[^rank_init]:
In reality, the rank values could also be any arbitrary value on the real number scale.

[^sum_or_avg]:
A sum of sums (instead of an average of sums) could be used here and the subsequent
results would not change.

[^re-scale]:
This is not completely necessary, but I believe that it makes the computation(s) and
the calculated values
"generalizable" for any `n`-length sequence.

Notably, even though the result set is split at the 50% threshold (which is a subjective choice),
this does not mean that 50% of all possible groups are classified as 50%.
(According to my method, only 33% are for `n = 4`.)

## Implementing the Heuristic

Ok, that is enough discussion. What follows is the implementation.

_NOTE:_ In order to keep focus on how the code implements methodology,
I recommend reviewing the code but not
worrying too much about the details (such as the internal workings
of my custom functions). Rather, I'd recommend inspecting in detail only the parts
that are printed out (and going back later to understand the complexities, if curious).

```{r packages_show, include = TRUE}
library("dplyr")
library("ggplot2")
library("tidyr")
# These packages are used, but their functions are called explicitly.
# library("purrr")
# library("broom")
# library("combinat")
```

The only
choice that I need to make to begin is the length of the set of values (i.e. `n`),
which should be a "small" integer.
I'll choose `4`, simply because `3` seems like it is "too small" and because subsequent
visualization(s) becomes "cluttered" (and interpretation becomes less direct)
if a number `5` or greater is chosen. [^trial_error] (Nonetheless, the methodology
and results remain valid for any integer.)

[^trial_error]:
I realized this through some trial and error.

```{r n}
n <- 4L
```

```{r n_combns}
n_combns <- factorial(n) * factorial(n)
n_rows <- n * n_combns
```


The following code chunk corresponds to steps 1 and 2 in my methodology, which are basically
just set-up steps.

```{r data_permns, include = TRUE, results = "hide"}
create_permns <- function(n = 1L) {
  
  n_seq <- seq(1L, n, by = 1L)
  
  combs <-
    combinat::permn(n_seq) %>%
    purrr::map( ~ paste(.x, collapse = "")) %>%
    unlist() %>%
    as.integer()
  data_xy <-
    tibble(grp_x = combs, grp_y = combs) %>%
    expand(grp_x, grp_y) %>%
    mutate(grp = paste0("x", grp_x, "y", grp_y))
  
  into_seq <- seq(1L, n, by = 1L) %>% as.character()
  sep_seq <- seq(1L, n - 1L, by = 1L)
  wrangle_data_xy <-
    function(data = NULL, which = NULL) {
      col_grp <- rlang::sym(paste0("grp_", which))
      col_0 <- rlang::sym(paste0(which, "0"))
      data %>%
        separate(!!col_grp, into_seq, sep = sep_seq, remove = FALSE) %>%
        gather(idx, !!col_0, matches("^[0-9]$")) %>%
        mutate_at(vars(idx, !!col_0), funs(as.integer))
    }
  inner_join(data_xy %>% wrangle_data_xy("x"),
               data_xy %>% wrangle_data_xy("y")) %>%
    select(-idx) %>%
    arrange(grp_x, grp_y)
}

data_permns <- create_permns(n = n)
data_permns
```

```{r data_permns_show, include = TRUE, echo = FALSE, results = "markup"}
data_permns %>%
  create_kable()
```

Note(s) about the above code chunk:

+ Although the "actual" ranks (`y0`) could be any real numbers,
I'm using values between `1` and `n` because my methodology dictates that the arbitrary
set of values would need to be transformed to this range anyways. When transformed to this
range, the values should be interpreted as ranks.

+ Like the "actual" `y0` ranks (representing the order of the original, arbitrary values),
the `x0` ranks could technically be any real numbers, but they
would need to be transformed to the `1`-to-`n` range anyways, so I do that directly.

+ The number of combinations of "actual" (`y0`) and "assigned" (`x0`) rank pairs is
equal to `n! * n!` (i.e. `r n_combns`).
For my implementation, the data.frame `data_permns` actually has 
`n! * n! * n` (`r n_rows`) rows (because it is arranged in a "long" format).

+ `grp_x` and `grp_y` (and the combination of the two in the `grp` column) 
identify the `n`-length groups of pairs of `x0` and `y0` ranks. These are primarily
useful for human interpretability and are not actually relevant for computations.

+ I use a function here (and elsewhere) although it may seem unnecessary
for a single execution because I'll repeat the methodology for different values of `n` later.

Now, I implement the initial calculation of "inverse monotonic distance" (`mntc_distinv`).

```{r data_mntc, include = TRUE, results = "hide"}
add_mntc_cols <- function(data = NULL) {
  data %>%
    group_by(grp) %>%
    arrange(x0, .by_group = TRUE) %>% 
    mutate(mntc = ifelse((y0 == cummax(y0)) | (y0 == cummin(y0)), 1L, 0L)) %>% 
    mutate(mntc_distinv = as.integer(x0 * (max(x0) - abs(x0 - y0)))) %>% 
    ungroup()
}
data_mntc <- add_mntc_cols(data_permns)
data_mntc
```

```{r data_mntc_show, include = TRUE, echo = FALSE, results = "markup"}
data_mntc %>% 
  mutate_if(is.numeric, funs(round(., 2))) %>% 
  create_kable()
```

```{r data_mntc_debug, eval = FALSE}
data_mntc %>% count(mntc, sort = TRUE)
data_mntc %>% count(mntc_distinv, sort = TRUE)
```

Note(s) about the above code chunk:

+ The `mntc` variable is a "running" binary `1` or `0` to indicate whether or not
`y0` is monotonic up through its position in the sequence. (It does not differentiate
between increasing or decreasing.)
+ I use the `integer` data type (as opposed to simply `numeric`) where possible
because it is more memory efficient (although memory efficiency is not a concern
with this data).


Next is the calculation of the transformed (i.e. summed-averaged-re-scaled) version 
of the "inverse monotonic distance" (`mntc_distinv`),
as well as the split of the `mntc_distinv` into upper and lower 50% quantiles (`mntc_tier2`).

```{r summ_mntc, include = TRUE, results = "hide"}
unitize <- function(x = NULL) {
  (x - min(x)) / (max(x) - min(x))
}
summarise_mntc <- function(data = NULL) {
  data %>%
    group_by(grp) %>% 
    summarise_at(vars(mntc_distinv), funs(mean)) %>% 
    ungroup() %>% 
    mutate_at(vars(mntc_distinv), funs(unitize)) %>% 
    mutate(mntc_tier2 = cut(mntc_distinv, 2))
}
summ_mntc <- summarise_mntc(data_mntc)
summ_mntc
```

```{r summ_mntc_show, include = TRUE, echo = FALSE, resuls = "markup"}
summ_mntc %>%
  mutate_if(is.numeric, funs(round(., 2))) %>% 
  create_kable()
```

```{r summ_mntc_debug, eval = FALSE}
summ_mntc %>% count(mntc_distinv, sort = TRUE)
summ_mntc %>% count(mntc_tier2, sort = TRUE)
# summ_mntc %>%
#   select(-matches("tier")) %>% 
#   gather(metric, value, starts_with("mntc")) %>% 
#   # group_by(metric) %>% 
#   # summarise_at(vars(value), funs(mean)) %>% 
#   # ungroup() %>% 
#   # count(mntc_distinv, sort = TRUE) %>%
#   ggplot(aes(x = value, fill = metric)) +
#   # geom_col()
#   geom_histogram() +
#   facet_wrap(~metric)
# summ_mntc %>%
#   select(matches("^mntc")) %>%
#   select_if(is.numeric) %>%
#   corrr::correlate()
```

Now, with the "algorithm" fully implemented, I can begin to evaluate the results.

Exactly how many values make up each 50% quantile?

```{r summ_mntc_cnts, include = TRUE, echo = FALSE, results = "markup"}
summ_mntc %>%
  count(mntc_tier2, sort = TRUE) %>%
  mutate(n_pct = 100 * (n / sum(n))) %>% 
  mutate_if(is.numeric, funs(round(., 2))) %>% 
  create_kable()
```
What does the distribution of all "inverse monotonic distance" values look like?

```{r viz_mntc_distinv, include = TRUE, echo = FALSE, fig.height = 6, fig.width = 8}
n_bins <- length(unique(summ_mntc$mntc_distinv))

lab_rank_seq <-
  paste0("Rank Combinations (", n, "-Length Sequence)")

lab_title_mntc_distinv <-
  paste0("Distribution of Inverse Monotonic Distance Values\nfor ", lab_rank_seq)
viz_mntc_distinv <-
  summ_mntc %>%
  ggplot(aes(x = mntc_distinv)) +
  geom_histogram(bins = 20) +
  theme_base() +
  labs(title = lab_title_mntc_distinv)
viz_mntc_distinv
```

```{r viz_summ_mntc, include = TRUE, echo = FALSE, fig.height = 10, fig.width = 10}
add_grps_xy <- function(data = NULL) {
  data %>%
    inner_join(data_permns %>% select(grp, grp_x, grp_y), by = "grp") %>% 
    mutate_at(vars(starts_with("grp_")), funs(as.character))
}
lab_title_summ_mntc <-
  paste0(
    lab_rank_seq,
    " Identified as\n\"Sufficiently\" Monotonic (Increasing)"
  )
lab_caption_prefix <-
  "Pairs identified as significant per "
lab_caption_summ_mntc <-
  paste0(lab_caption_prefix, "custom methodology marked with dot ('.').")

viz_summ_mntc <-
  summ_mntc %>%
  add_grps_xy() %>% 
  # mutate(lab = round(mntc_mean, 1)) %>% 
  ggplot(aes(x = grp_x, y = grp_y)) +
  geom_tile(aes(fill = mntc_distinv)) +
  scale_fill_gradient2(low = "purple", high = "yellow", mid = "grey95", midpoint = 0.5) +
  geom_point(
    data =
      summ_mntc %>%
      filter(mntc_tier2 == "(0.5,1]") %>%
      add_grps_xy(),
    shape = 20, size = 1
  ) +
  # geom_text(aes(label = lab), color = "black", fontface = "bold") +
  coord_equal() +
  theme_base() +
  theme_tile() +
  guides(fill = guide_legend(title = "Weighted\nMonotocity")) +
  theme(legend.position = "right") +
  labs_tile() +
  labs(
    title = lab_title_summ_mntc,
    caption = lab_caption_summ_mntc
  )
viz_summ_mntc
```

The positive identificaiton (in yellow) of combinations along the left-to-right, lower-to-upper
diagonal is what I would expect. These are the values where `x0` and `y0` are perfectly matched. Conversely, values along the other diagonal are generally
colored in purple, as I would expect. These combinations consist of sequences of `x0` and `y0` values
that are "negatively" symmetric (e.g. (`1`, `2`, `3,` `4`) and (`4`, `3`, `2`, `1`)).

## Checking the Heuristic

Ok, my heuristic seems valid, but how can I know for sure that it is reasonable?
I mentioned before that Spearman's rho should serve a good measure, so I'll take
a look at it now.

```{r summ_mntc_wcortest, include = TRUE, results = "hide"}
add_cortest_cols <- function(data = NULL) {
  data %>%
    group_by(grp) %>%
    nest() %>%
    mutate(cortest =
             purrr::map(data, ~ broom::tidy(cor.test(.$x0, .$y0, method = "spearman")))
    ) %>%
    unnest(cortest, .drop = TRUE) %>%
    select(grp, estimate, p.value)
}
summarise_mntc_wcortest <- function(data = NULL) {
   summ <- summarise_mntc(data)
   data %>%
     add_cortest_cols() %>% 
     inner_join(summ, by = "grp")
}

summ_mntc_wcortest <- summarise_mntc_wcortest(data_mntc)
summ_mntc_wcortest
```

```{r summ_mntc_wcortest_show, include = TRUE, echo = FALSE, results = "markup"}
summ_mntc_wcortest %>%
  mutate(n_pct = 100 * (n / sum(n))) %>% 
  mutate_if(is.numeric, funs(round(., 2))) %>% 
  create_kable()
```


```{r summ_mntc_wcortest_alt, eval = FALSE}
# # Alternatively... 
# data_mntc %>%
#   group_by(grp) %>% 
#   summarise(
#     cortest = cor.test(x0, y0)$estimate,
#     p.value = cor.test(x0, y0)$p.value
#   ) %>% 
#   ungroup()
# # Or...
# data_mntc %>%
#   group_by(grp) %>% 
#   do(broom::tidy(cor.test(.$x0, .$y0, method = "spearman"))) %>% 
#   ungroup()
```


What exactly is the distribution of the Pearson's rho t-test estimates and p-values?

```{r summ_mntc_joined_cnts, include = TRUE, echo = FALSE, results = "markup"}
summ_mntc_wcortest %>%
  mutate_if(is.numeric, funs(round(., 2))) %>% 
  count(abs(estimate), p.value, sort = TRUE) %>% 
  create_kable()
```

Note(s) about the above output:

+ By taking the absolute value of the estimate, I am essentially treating
monotonically increasing and decreasing as equal.
+ There are a relatively "small" number of distinct values. [^small_distinct]

[^small_distinct]:
Here, there are only `n + 1` (i.e. `r n + 1` unique `abs(estimate)`s and `p.value`s.
This result is __not__ generally true.
(For example, when choosing `n = 5`, there will be more than `6` unique values of each metric.)

Now, to understand how the Pearson's rho t-test `estimate`s and `p.value`s correspond to my heuristic,
I'll simply overlay the combinations that are identified as significant to my
previous heat map of rank combinations. Because I'm erring on the side of flexibility
in defining "sufficient" monotonicity, I'll say that the pairs corresponding to the
bottom two tiers of p-values
(corresponding to `0.0833` and `0.33`) constitute "sufficient" monoticity. 

```{r viz_summ_mntc_wcortest, include = TRUE, echo = FALSE, fig.height = 10, fig.width = 10}
lab_caption_summ_mntc_wcortest <-
  paste0(
    lab_caption_summ_mntc, "\n",
    lab_caption_prefix, "Pearson's rho p.value < 0.33 marked with 'x'.\n",
    lab_caption_prefix, "Pearson's rho p.value < 0.083 marked with asterisk ('*')."
  )
viz_summ_mntc_wcortest <-
  viz_summ_mntc +
  geom_point(
    data =
      summ_mntc_wcortest %>% 
      add_grps_xy() %>% 
      filter(estimate > 0) %>% 
      filter(p.value < 0.4),
    shape = 4, size = 3
  ) +
  geom_point(
    data =
      summ_mntc_wcortest %>% 
      add_grps_xy() %>% 
      filter(estimate > 0) %>% 
      filter(p.value < 0.1),
    shape = 8, size = 3
  ) +
  labs(caption = lab_caption_summ_mntc_wcortest)
viz_summ_mntc_wcortest
```

```{r viz_summ_mntc_wcortest_export, eval = FALSE}
teproj::export_ext_png(
  viz_summ_mntc_wcortest,
  dir = params$dir_viz,
  export = params$export_viz,
  units = "in",
  width = 10,
  height = 10
)
```

It looks like there is a large amount of overlap between my heuristic
classification of "sufficient" monoticity and that identified by a more statistical 
approach.

Now I'll repeat the simulation for other values of `n`.
(Because computations start to become intensive with `n = 6`, and because
the `n = 2` is relatvily trivial, I'll evaluate values of `3`, `4`, and `5` for `n`. [^map-nest])

[^map-nest]:
This presents a good opportunity to implement a version of the "nest-mutate-unnest" idiom
that can be very effective for creating many models. The
["many models" chapter](http://r4ds.had.co.nz/many-models.html) in the
[_R For Data Science_ book](http://r4ds.had.co.nz) provides an excellent example of this process.


```{r summ_mntc_byn, include = TRUE, echo = TRUE, eval = FALSE}
ns <- tibble(n = 3L:5L)
summ_mntc_byn <-
  ns %>% 
  mutate(data = purrr::map(n, ~(create_permns(.x) %>% add_mntc_cols()))) %>% 
  mutate(summ = purrr::map(data, summarise_mntc_wcortest)) %>% 
  unnest(summ, .drop = TRUE) %>% 
  ungroup() %>% 
  arrange(n)
```

```{r summ_mntc_byn_export, eval = FALSE}
readr::write_csv(summ_mntc_byn, file.path(params$dir_data, "summ_mntc_byn.csv"))
```
```{r summ_mntc_byn_import, include = TRUE, echo = FALSE}
path_data <- "https://raw.githubusercontent.com/tonyelhabr/rank-regression/master/data/summ_mntc_byn.csv"
# summ_mntc_byn <- readr::read_csv(file.path(params$dir_data, "summ_mntc_byn.csv"))
summ_mntc_byn <- readr::read_csv(path_data)
```

```{r summ_mntc_byn_debug, eval = FALSE}
summ_mntc_byn %>% 
  filter(n == 5) %>% 
  ggplot(aes(x = estimate)) +
  geom_density()

summ_mntc_byn %>%
  filter(n == 5) %>% 
  mutate(estimate_abs = abs(estimate)) %>% 
  count(n, estimate_abs >= 0.5, sort = TRUE) %>% 
  group_by(n) %>% 
  mutate(nn_pct = 100 * (nn / sum(nn))) %>% 
  ungroup()
summ_mntc_byn %>%
  filter(n == 5) %>% 
  mutate(estimate_abs = abs(estimate)) %>% 
  filter(estimate_abs >= 0.8) %>% 
  count(n, estimate_abs, p.value, sort = TRUE) %>% 
  group_by(n) %>% 
  mutate(nn_pct = 100 * (nn / sum(nn))) %>% 
  ungroup()

summ_mntc_byn %>% 
  filter(n == 5) %>% 
  mutate(estimate_abs = abs(estimate)) %>% 
  select(mntc_distinv, estimate_abs, p.value) %>%
  corrr::correlate()

summ_mntc_byn %>% 
  select(-mntc_tier2) %>% 
  gather(metric, value, -n, -grp) %>% 
  widyr::pairwise_cor(
    item = metric,
    feature = n,
    value = value,
    method = "spearman",
    use = "everything"
  )
```

What is the breakdown of `mntc_tier2` values?

```{r summ_mntc_byn_show, include = TRUE, echo = FALSE, results = "markup"}
add_n_pct_col <- function(data = NULL) {
  data %>% 
    group_by(n) %>% 
    mutate(nn_pct = 100 * (nn / sum(nn))) %>% 
    ungroup()
}

summ_mntc_byn %>% 
  count(n, mntc_tier2) %>% 
  add_n_pct_col() %>% 
  mutate_if(is.numeric, funs(round(., 1))) %>% 
  create_kable()
```

What about the distribution of `mntc_distinv` values? And of the `estimate`s and `p.value`s?

```{r summ_mntc_byn_show_2, include = TRUE, echo = FALSE, results = "markup"}
summ_mntc_byn %>% 
  count(n, abs(estimate), p.value, mntc_tier2) %>% 
  add_n_pct_col() %>% 
  arrange(n, mntc_tier2) %>% 
  mutate_if(is.numeric, funs(round(., 1))) %>% 
  create_kable(20)
```

```{r viz_mntc_distinv_byn, include = TRUE, echo = FALSE, fig.height = 6, fig.width = 8}
lab_title_mntc_distinv_byn <-
  gsub("\\s\\(.*$", ", Different n-Length Sequences", lab_title_mntc_distinv)

viz_mntc_distinv_byn <-
  summ_mntc_byn %>% 
  mutate(n = paste0("n = ", n)) %>% 
  ggplot(aes(x = mntc_distinv)) +
  geom_histogram(bins = 20, color = "black") +
  theme_base() +
  theme(panel.background = element_rect()) +
  facet_wrap( ~ n, scales = "free") +
  labs(title = lab_title_mntc_distinv_byn)
viz_mntc_distinv_byn
```

```{r viz_mntc_distinv_byn_export, eval = FALSE}
teproj::export_ext_png(
  viz_mntc_distinv_byn,
  dir = params$dir_viz,
  export = params$export_viz,
  units = "in",
  width = 8,
  height = 6
)
```

```{r viz_mntc_tier2_byn, include = TRUE, echo = FALSE, fig.height = 8, fig.width = 6}
lab_title_mntc_tier2_byn <-
  gsub("^.*,", "Distribution of Pearson's Rho T-Test\nEstimates and P-Values,", lab_title_mntc_distinv_byn)
lab_subtitle_mntc_tier2_byn <-
  "Colored According to Custom Heuristic Classification"
  # NULL
lab_caption_mntc_tier2_byn <-
  paste0(
    "Estimate expectation: yellow more dense near 1 (right); purple mre dense near -1 (left).\n",
    "P-value expectation: yellow more dense near 0 (left); purple more dense near 1 (right).\n",
    "Caveat: Small sample size (i.e. lack of unique p-values)."
  )

viz_mntc_tier2_byn <-
  summ_mntc_byn %>%
  # mutate_at(vars(p.value), funs(if_else(. >= 1, 0.9, .))) %>% 
  mutate(n = paste0("n = ", n)) %>% 
  gather(metric, value, estimate, p.value) %>% 
  # count(n, p.value, mntc_tier2) %>% 
  ggplot(aes(x = value, fill = mntc_tier2)) +
  geom_histogram(bins = 25, color = "black") +
  # geom_bar(aes(y = ..count..), color = "black") +
  # geom_col(aes(y = ..density..)) +
  scale_fill_manual(values = c("purple", "yellow")) +
  facet_grid(n ~ metric, scales = "free") +
  theme_base() +
  theme(panel.background = element_rect()) +
  # xlim(-1, 1) +
  labs(
    title = lab_title_mntc_tier2_byn,
    subtitle = lab_subtitle_mntc_tier2_byn,
    caption = lab_caption_mntc_tier2_byn
  )
viz_mntc_tier2_byn
```

```{r viz_mntc_tier2_byn_export, eval = FALSE}
teproj::export_ext_png(
  viz_mntc_tier2_byn,
  dir = params$dir_viz,
  export = params$export_viz,
  units = "in",
  width = 6,
  height = 8
)
```


The distributions are sparse due to the relatively small number of unique
values for each metric (`mntc_distinv`, `p.value`, etc.). [^sparse]
Consequently, it is a bit difficult to extract
much meaningful insight about the relationships among the metrics. To really understand
how the distributions and relationships scale with larger values of `n`, mathematical
theory would need to be applied.

[^sparse]:
Unfortunately this
is due to the nature of the data and the simulation, so nothing can be done
about it.

Nonethless, without jumping more into statistical theory,
It seems to me that the identification of rank combinations as significant by my heuristic classification
and Spearman's rho (assuming that one uses the traditional p-value-below-a-thresshold approach)
would become __more dissimilar__ as the value of `n` increases. This is
because my classification simply splits all possible values into two sets for any 
value of `n`, meaning that the percentage of all possible combinations is
relatively insensitive to the value of `n`. [^n4_result] On the other hand,
the _Spearman's rho_ p-values would become more refined with larger values of `n`.

[^n4_result]:
Note that the 33% number found for `n = 4`
is not generally true, although this percentage does not seem to change drastically
with different values of `n`.

Anyways, I believe that my heuristic serves my purposes well.
I only really intended it to be used for small values of `n`. Also, I intended to create
a "relaxed" definition of monotonocity, so having only a very small percentage of all possible
rank combinations meet the criteria would have actually been undesireable.

```{r save_session}
save.image(file = file.path(params$dir_data, "rank-regression-1.RData"))
```

## Conclusion

In the end, I think I did more work than
I really needed to do to answer my original question about
quantifying monotonocity and inferring significance,
but I think, in all, this was a worthwhile exploration.

