---
title: Analyzing Professional Sports Team Colors with R, Part 2
slug: sports-colors-analysis-r-part-2
date: "2018-04-01"
categories: []
tags:
 - r
 - sports
 - nba
 - colors
 - ggplot2
banner: "img/sports-colors-analysis-r-part-2/viz_colors_ord2_nba_rgb_dist.png"
---


```{r setup, echo = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  echo = TRUE,
  # echo = FALSE,
  cache = TRUE,
  # results = "markdown",
  # results = "hide",
  fig.align = "center",
  # fig.show = "asis",
  fig.width = 10,
  fig.height = 6,
  # out.width = 6,
  # out.height = 6,
  warning = FALSE,
  message = FALSE
)

```

_NOTE: This write-up picks up where [the previous one](https://tonyelhabr.rbind.io/post/sports-colors-analysis-r/) left off. All of the session data is carried over._

```{r restore_session, echo = FALSE, eval = TRUE}
# session::restore.session(file.path("static", "data", "colors-eda-1.RData"))
# load(file.path("static", "data", "colors-eda-1.RData"))
url <- "https://raw.githubusercontent.com/tonyelhabr/colors-eda/master/data/colors-eda-1.RData"
load(url(url))
# vars_all <- ls()
# vars_keep <- stringr::str_subset(vars_all, "^add_|^convert_|^lab_|^viz_*_base$|(colors_tidy_ord2)")
# rm(list = setdiff(vars_all, vars_keep))
library("dplyr")
library("ggplot2")
viz_theme_base <-
  teplot::theme_te() +
  theme(
    legend.position = "none",
    # panel.background = element_rect(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.ticks = element_blank(),
    plot.caption = element_text(face = "plain"),
    axis.text.x = element_text(angle = 90)
  )
create_kable <- function(data, num_show = 10, full_width = TRUE) {
  
  num_rows <- nrow(data)
  show_fn <- ifelse(num_rows > num_show, TRUE, FALSE)
  if(show_fn) {
    data <- data %>% slice(1:num_show)
  }
  
  out <-
    data %>% 
    knitr::kable( "html", escape = FALSE) %>%
    kableExtra::kable_styling(full_width = full_width, position = "center")
  
  if(show_fn) {
    out <-
      out %>%
      kableExtra::add_footnote(c(sprintf("# of total rows: %.0f", num_rows)), notation = "number")
  }
  out
}
```

## Color Similarity

Now, I'd like to evaluate color similarity more closely. To help verify 
any quantitative deductions with some intuition,
I'll consider only a single league for this--the NBA, the league that I know the best. 

Because I'll end up plotting team names at some point and
some of the full names are relatively lengthy, I want
to get the official abbreviations for each team.
Unfortunately, these
don't come with the [`teamcolor` package](https://github.com/beanumber/teamcolors),
but I can use [Alex Bresler](http://asbcllc.com/)'s 
[`nbastatR` package](https://github.com/abresler/nbastatR) to get them.

```{r nms_nba}
# Assign `df_dict_nba_teams` to Global environment.
nbastatR::assign_nba_teams()
nms_nba <-
  teamcolors::teamcolors %>% 
  filter(league == "nba") %>% 
  inner_join(
    df_dict_nba_teams %>%
      setNames(snakecase::to_snake_case(names(.))) %>%
      filter(!is_non_nba_team) %>% 
      select(name = name_team, slug = slug_team),
    by = c("name")
  )

colors_tidy_ord2_nba <-
  nms_nba %>% 
  select(name, league, slug) %>% 
  inner_join(colors_tidy_ord2, by = c("name", "league"))
```

To give the unfamiliar reader a better understanding of what exactly this subset of the
`teamcolors` data incorporate, here's a 
visualization of the primary and secondary colors of all NBA teams

```{r viz_colors_ord2_nba, echo = FALSE, fig.height = 8, fig.width = 8}
colors_ord2_nba <-
  colors_tidy_ord2_nba %>% 
  tidyr::spread(ord, hex)

lab_title_colors_ord2_nba <-
  stringr::str_replace_all(lab_subtitle_kclusts, lab_lgs, "NBA")

# Reference:
# + https://github.com/beanumber/teamcolors.
viz_colors_ord2_nba <-
  colors_ord2_nba  %>% 
  mutate(name_slug = paste0(name, ": ", slug)) %>% 
  ggplot(aes(x = name_slug, color = name_slug, fill = name_slug)) + 
  geom_bar(size = 2) +
  coord_flip() +
  scale_x_discrete(NULL) + 
  scale_y_continuous(NULL) + 
  scale_fill_manual(values = colors_ord2_nba$primary) + 
  scale_color_manual(values = colors_ord2_nba$secondary) + 
  viz_theme_base +
  theme(axis.text.x = element_blank(), 
        axis.ticks.x = element_blank()) +
  viz_labs_base +
  labs(title = lab_title_colors_ord2_nba)
viz_colors_ord2_nba
```


After grabbing the abbreviations (or `slug`s), I can move on to breaking up
the hex values into their RGB components. [^fn_rgb]
I'll be looking at only the primary and secondary colors again.

[^fn_rgb]:
I use the same method as the one I used before.

```{r colors_ord2_nba_rgb_tidy}
colors_ord2_nba_rgb_tidy <-
  colors_tidy_ord2_nba %>%
  add_rgb_cols() %>% 
  select(-hex) %>%
  tidyr::gather(rgb, value, red, green, blue)

colors_ord2_nba_rgb_tidy %>% 
  create_kable()
```

```{r ggimage, include = FALSE, eval = FALSE}
# Reference:
# + https://cran.r-project.org/web/packages/ggimage/vignettes/ggimage.html.
# n <- 3
# path_img <- file.path("data-raw", "sas.png")
# path_img <- "http://i.cdn.turner.com/nba/nba/.element/media/2.0/teamsites/spurs/Schad/img/PostgameTeams/SAS.png"
# path_img_dwnld <- "https://github.com/jimniels/teamcolors/blob/master/static/img/nfl/arizona-cardinals.svg"
# convertGraph::convertGraph(path_img_dwnld, "out.png")
# # path_img <- "content.sportslogos.net/logos/7/177/full/7420_arizona_cardinals-primary_on_dark-2005.png"
# set.seed(42)
# d <- data.frame(
#   x = rnorm(n),
#   y = rnorm(n),
#   img = path_img, 
#   size = n
# )
# d %>% 
#   ggplot(aes(x, y)) + 
#   ggimage::geom_image(aes(image = img), size = .05)

```

With the RGB values extracted, I can use the `widyr::pairwise_dist()` function
to compute the relative distance among teams in terms of RGB values for each color 
ordinality.I think the default 
method--["Euclidean" distance](https://en.wikipedia.org/wiki/Euclidean_distance)--is reasonable.

```{r do_pairwise_dist}
do_pairwise_dist <- function(data, method) {
  data %>% 
    group_by(ord) %>% 
    widyr::pairwise_dist(name, rgb, value, upper = TRUE, method = method) %>% 
    rename(name1 = item1, name2 = item2) %>% 
    select(everything(), value = ncol(.)) %>% 
    arrange(value, .by_group = TRUE) %>% 
    ungroup()
}

```

As one might expect, there's not much difference between these 
two distance methods (if correlation is deemed a valid metric for quantifying similarity).

How exactly do all of the individual distances compare?

```{r colors_ord2_nba_rgb_dist, include = FALSE}
colors_ord2_nba_rgb_dist <-
  colors_ord2_nba_rgb_tidy %>%
    do_pairwise_dist("euclidean")

colors_ord2_nba_rgb_dist_ranks <-
  colors_ord2_nba_rgb_dist %>%
  group_by(ord, name1) %>%
  summarize(value = mean(value)) %>%
  ungroup() %>%
  group_by(name1) %>% 
  summarize(value = mean(value)) %>% 
  ungroup() %>% 
  mutate(rank = row_number(value)) %>% 
  arrange(rank)
```

```{r colors_ord2_nba_rgb_dist_debug, include = FALSE, eval = FALSE}
colors_ord2_nba_rgb_dist %>% 
  group_by(ord, name1) %>% 
  do(slice(., c(1, n()))) %>% 
  ungroup()
colors_ord2_nba_rgb_dist %>% 
  arrange(value) %>%
  filter(ord == "primary") %>% 
  filter(value != 0) %>% 
  slice(c(1:10, c(n()-9):n())) %>% 
  ungroup()
colors_ord2_nba_rgb_tidy %>%
  filter(name %in% c("San Antonio Spurs", "Pittsburgh Steelers", "Brooklyn Nets")) %>%
  tidyr::spread(rgb, value)
```


```{r viz_colors_ord2_nba_rgb_dist, echo = FALSE, fig.height = 10, fig.width = 6}
lvls_ord2_nba_rgb_dist_ranks <-
  colors_ord2_nba_rgb_dist_ranks %>%
  pull(name1)

labs_ord2_nba_rgb_dist_ranks <-
  colors_ord2_nba_rgb_dist_ranks %>%
  left_join(nms_nba %>% select(name1 = name, slug1 = slug, league1 = league), by = c("name1")) %>% 
  pull(slug1)

lab_title_colors_ord2_nba_rgb_dist <-
  "Pairwise Distance of RGB Values"
lab_subtitle_colors_ord2_nba_rgb_dist <-
  lab_title_colors_ord2_nba
lab_caption_colors_ord2_nba_rgb_dist_base <-
  paste0(
    "Similarity is based on Euclidean distance of RGB values.\n",
    "Order is determined by overall similarity across both primary and secondary colors."
  )
lab_caption_colors_ord2_nba_rgb_dist <-
  paste0(
    lab_caption_colors_ord2_nba_rgb_dist_base,
    "\nAverage is shown in team colors."
  )

viz_colors_ord2_nba_rgb_dist <-
  colors_ord2_nba_rgb_dist %>% 
  mutate(name1 = factor(name1, levels = lvls_ord2_nba_rgb_dist_ranks)) %>% 
  ggplot(aes(x = name1, y = value)) +
  geom_point(alpha = 0.25, color = "grey50") +
  geom_point(
    data = colors_ord2_nba_rgb_dist_ranks,
    aes(color = name1, fill = name1),
    shape = 21,
    size = 5
  ) +
  scale_color_manual(values = colors_ord2_nba$secondary) +
  scale_fill_manual(values = colors_ord2_nba$primary) +
  scale_x_discrete(labels = labs_ord2_nba_rgb_dist_ranks) +
  coord_polar(theta = "x") +
  viz_theme_base +
  theme(
    axis.ticks = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_text(angle = 0)
  ) +
  viz_labs_base +
  labs(
    title = lab_title_colors_ord2_nba_rgb_dist,
    subtitle = lab_subtitle_colors_ord2_nba_rgb_dist,
    caption = lab_caption_colors_ord2_nba_rgb_dist
  )
viz_colors_ord2_nba_rgb_dist
```


I think that the above plot does a good job of
highlighting the average distance values (in terms of RGB)
of each team. Additionally, by sorting the teams by value, it illustrates
exactly which teams are the most "generic" (i.e. most similar to all other teams)
and the most "unique" (i.e. least similar to all other teams.)

I can also use a heat map to visualize the same data
(Who doesn't like a good heat map?)

```{r colors_ord2_nba_rgb_dist_byord_viz, echo = FALSE}
create_colors_ord2_nba_rgb_dist_viz <- function(data, ranks) {
  data %>% 
    # tidyr::complete(name1, name2) %>% 
    inner_join(
      ranks %>% select(name1, rank1 = rank),
      by = c("name1")
    ) %>% 
    inner_join(
      ranks %>% select(name2 = name1, rank2 = rank),
      by = c("name2")
    ) %>% 
    # mutate_at(vars(name1, name2), funs(factor(., levels = lvls_ord2_nba_rgb_dist_ranks)))
    mutate(name1 = reorder(name1, rank1),
           name2 = reorder(name2, rank2))
}

colors_ord2_nba_rgb_dist_byord_viz <-
  colors_ord2_nba_rgb_dist  %>%
  create_colors_ord2_nba_rgb_dist_viz(colors_ord2_nba_rgb_dist_ranks)
```

```{r viz_colors_ord2_nba_rgb_dist_setup, echo = FALSE}
color_top <- "green"
color_bot <- "red"
lab_caption_colors_ord2_nba_rgb_dist_2 <-
  paste0(
    lab_caption_colors_ord2_nba_rgb_dist_base,
    "\nDark = similar; light = dissimilar.",
    "\nTop and bottom 2% most similar team-team colors emphasized with ",
    color_top,
    " and ",
    color_bot,
    "."
  )
viz_labs_colors_ord2_nba_rgb_dist <-
  labs(
    title = lab_title_colors_ord2_nba_rgb_dist,
    subtitle = lab_subtitle_colors_ord2_nba_rgb_dist,
    caption = lab_caption_colors_ord2_nba_rgb_dist_2,
    x = NULL,
    y = NULL
  )

visualize_colors_ord2_nba_rgb_dist_const <-
  function(data, value = "value", emphasize_top = TRUE, pct_top = 0.02) {

    value_sym <- rlang::sym(value)
    viz <-
      data %>% 
      # filter(league1 == league2) %>%
      # mutate(lg_ord = paste0(toupper(league1), ", ", stringr::str_to_title(ord))) %>% 
      mutate(!!value_sym := ifelse(is.na(!!value_sym), 0, !!value_sym)) %>% 
      ggplot(aes(x = name1, y = name2)) +
      geom_tile(aes_string(fill = value_sym)) +
      scale_x_discrete(labels = labs_ord2_nba_rgb_dist_ranks) +
      scale_y_discrete(labels = labs_ord2_nba_rgb_dist_ranks) +
      viridis::scale_fill_viridis(option = "E") +
      coord_fixed() +
      # scale_fill_gradient(low = "black", high = "white") +
      viz_theme_base +
      theme(axis.text.x = element_text(angle = 90)) +
      viz_labs_colors_ord2_nba_rgb_dist
    
    if(emphasize_top) {
      data_pranked <-
        data %>% 
        mutate(prank = percent_rank(!!value_sym))
      viz <-
        viz +
        geom_tile(
          data = data_pranked %>%
            filter(prank <= pct_top),
          fill = color_top
        ) +
        geom_tile(
          data = data_pranked %>%
            filter(prank >= (1 - pct_top)),
          fill = color_bot
        )
    }
    viz
  }
```

```{r viz_colors_ord2_nba_rgb_dist_byord, include = FALSE, eval = FALSE}
# viz_colors_ord2_nba_rgb_dist_byord <-
#   colors_ord2_nba_rgb_dist_byord_viz %>%
#   visualize_colors_ord2_nba_rgb_dist_const() +
#   facet_wrap(~ ord, scales = "free")
# viz_colors_ord2_nba_rgb_dist_byord
```


```{r colors_ord2_nba_rgb_dist_viz, include = FALSE}
colors_ord2_nba_rgb_dist_viz <-
  colors_ord2_nba_rgb_dist %>%
  group_by(name1, name2) %>% 
  summarize(value = sum(value)) %>% 
  ungroup() %>% 
  create_colors_ord2_nba_rgb_dist_viz(colors_ord2_nba_rgb_dist_ranks)
```

```{r viz_colors_ord2_nba_rgb_dist_2, echo = FALSE, fig.height = 8, fig.width = 8}
viz_colors_ord2_nba_rgb_dist_2 <-
  colors_ord2_nba_rgb_dist_viz %>%
  visualize_colors_ord2_nba_rgb_dist_const()
viz_colors_ord2_nba_rgb_dist_2
```

Like with the previous plot, I order the teams on each axis
by total distance from all other teams--teams with the highest cumulative similarity to all
other teams appear towards the bottom and left, while teams that
contrast most with all others appear towards the top and right. And,
to add some nuance, I emphasize the individual pairs that have the highest
and lowest similarity with different colors.

Exactly which teams match most and least closely with one another (in terms of color similarity)?
Here's a list of the top and bottom matches for each team.

```{r colors_ord2_nba_rgb_dist_viz_prank, echo = FALSE, eval = FALSE}
# colors_ord2_nba_rgb_dist_viz_prank <-
#   colors_ord2_nba_rgb_dist_viz %>% 
#   mutate(prank = percent_rank(value)) %>% 
#   arrange(prank)
# 
# slice_topbot <- function(data, num) {
#   data %>%
#     slice(c(1:num, c(n() - (num - 1)):n()))
# }
# 
# colors_ord2_nba_rgb_dist_viz_prank %>% 
#   select(-rank1, -rank2, -prank) %>% 
#   slice_topbot(10) %>% 
#   create_kable(num_show = 20)
```

```{r colors_ord2_nba_rgb_dist_viz_topbot, echo = FALSE}
colors_ord2_nba_rgb_dist_viz %>% 
  group_by(name1) %>% 
  arrange(value, .by_group = TRUE) %>% 
  do(slice(., c(1, n()))) %>% 
  ungroup() %>%
  select(rank_overall = rank1, everything(), -rank2) %>% 
  rename(dist = value) %>%
  mutate_at(vars(dist), funs(round(., 0))) %>% 
  create_kable(num_show = 60)
```

These results don't really agree with what I--and maybe other NBA fans--would have guessed.
The Sacramento Kings (`SAC`)
have purple as their primary color, which is relatively unusual. I would think that they
would be in the lower half of these rankings. Whats going on? ...

## Color Theory

When doing this color-based analysis, several questions came to mind:

1. Is the [RGB model](https://en.wikipedia.org/wiki/RGB_color_model) 
really the best framework to use for comparing colors? What about the 
[HSL](https://en.wikipedia.org/wiki/HSL_and_HSV) (Hue, Saturation, Lightness) model?
Additionally, a quick Google search for
["What is the best method for identifying similarity between colors?"](https://stackoverflow.com/questions/5392061/algorithm-to-check-similarity-of-colors)
indicates the [YUV representation](https://en.wikipedia.org/wiki/YUV)--a model
I hadn't heard of before--is best,
(if human perception is the main concern).

2. Is Euclidean distance the best "distance" method to use?
But, because I'm curious, I'll look at how different the results would be if the
["Manhattan" distance](https://en.wiktionary.org/wiki/Manhattan_distance) is used instead.

3. Is "distance" even the best method for determining color similarity.
Why not a "similarity" metric (such as [cosine similarity](https://en.wikipedia.org/wiki/Cosine_similarity))?

Since I'm not expert in color models, and because I there is no definitive/conclusive
research with which I can cross-check my findings for color similarity among NBA teams,
I think its worthwhile to explore these questions in more detail.
First, I'll need to create HSL and YUV variations of the color data that I can
compare to the RGB version that I've used up to this point. (This
will help me answer the first question.) [^fn_hsl_yuv_code]
Then, with each of these data sets in hand, I'll tackle the latter two questions directly.
In the end, by comparing the different models with different methods,
I hope to come to some stronger conclusions and/or justifications of my findings
about NBA team colors.

[^fn_hsl_yuv_code]:
I don't show the code for this, so check out the .Rmd document for detail.

```{r yuv_funcs, echo = FALSE}
convert_rgb2yuv <- function(r, g, b) {
  out <- colorscience::RGB2YUV(c(r, g, b))
  c(y = out[1], u = out[2], v = out[3])
}

convert_hex2yuv <- function(hex, unname = FALSE) {
  rgb <- t(grDevices::col2rgb(hex))
  out <- convert_rgb2yuv(rgb[1], rgb[2], rgb[3])
  if(unname) {
    out <- unname(out)
  }
  out
}

add_yuv_cols <- function(data) {
  data %>%
    pull(hex) %>%
    purrr::map(convert_hex2yuv) %>% 
    purrr::reduce(rbind) %>% 
    tibble::as_tibble() %>%
    bind_cols(data, .)
}
```

```{r colors_ord2_nba_hsl_yuv_tidy, echo = FALSE}
colors_ord2_nba_hsl_tidy <-
  colors_tidy_ord2_nba %>%
  add_hsl_cols() %>% 
  select(-hex) %>%
  tidyr::gather(rgb, value, h, s, l)

colors_ord2_nba_yuv_tidy <-
  colors_tidy_ord2_nba %>%
  add_yuv_cols() %>% 
  select(-hex) %>%
  tidyr::gather(rgb, value, y, u, v)
```

### Euclidean Distance vs. Manhattan Distance

I'll look at two distance methods--Euclidean and Manhattan--to justify
my choice of Euclidean distance before.
To do this, I want to verify that the similarity determined by the
two methods is nearly identical.
(I would be surprised if they aren't.)

```{r colors_ord2_nba_dist, echo = FALSE}
correlate_pairwise_out2 <- function(data1, data2) {
  corrr::correlate(
    data1 %>%
      rename(value1 = value) %>%
      left_join(data2 %>% rename(value2 = value)) %>%
      select_if(is.numeric)
  )
}

rename_pairwise_dist_out <- function(data, model = c("rgb", "hsl", "yuv")) {
  nm1 <- rlang::sym(paste0(model, "_", "euclidean"))
  nm2 <- rlang::sym(paste0(model, "_", "manhattan"))
  data %>%
    rename(!!nm1 := value1, !!nm2 := value2) %>% 
    # mutate(rowname = ifelse(rowname == "value1", !!nm1, !!nm2))
    select(-rowname)
}

correlate_pairwise_out2(
  colors_ord2_nba_rgb_tidy %>%
    do_pairwise_dist("euclidean"),
  colors_ord2_nba_rgb_tidy %>%
    do_pairwise_dist("manhattan") 
) %>% 
  rename_pairwise_dist_out("rgb") %>% 
  mutate_if(is.numeric, funs(convert_dec2pct)) %>% 
  create_kable()


correlate_pairwise_out2(
  colors_ord2_nba_hsl_tidy %>%
    do_pairwise_dist("euclidean"),
  colors_ord2_nba_hsl_tidy %>%
    do_pairwise_dist("manhattan")
) %>% 
  rename_pairwise_dist_out("hsl") %>% 
  mutate_if(is.numeric, funs(convert_dec2pct)) %>% 
  create_kable()

correlate_pairwise_out2(
  colors_ord2_nba_yuv_tidy %>%
    do_pairwise_dist("euclidean"),
  colors_ord2_nba_yuv_tidy %>%
    do_pairwise_dist("manhattan")
) %>% 
  rename_pairwise_dist_out("yuv") %>% 
  mutate_if(is.numeric, funs(convert_dec2pct)) %>% 
  create_kable()
```

Indeed, it looks like there is high correlation found between the Euclidean and Manhattan
distances calculated when the hex color values are broken down into color components,
regardless of whether the RGB, HSL, or YUV representation is used.

Now, when keeping the distance method constant (Euclidean),
how do the color models compare?

```{r colors_ord2_nba_rgb_hsl_yuv_dist, echo = FALSE}
cor_dist_data <-
  colors_ord2_nba_rgb_tidy %>% 
  do_pairwise_dist("euclidean") %>% 
  rename(rgb_dist = value) %>% 
  left_join(
    colors_ord2_nba_hsl_tidy %>% 
      do_pairwise_dist("euclidean") %>% 
      rename(hsl_dist = value)
  ) %>%
  left_join(
    colors_ord2_nba_yuv_tidy %>% 
      do_pairwise_dist("euclidean") %>% 
      rename(yuv_dist = value)
  )
# A "tidy" method (i.e. not using `stats::cor()` that doesn't produce a plot.
cor_dist_data %>%
  select_if(is.numeric) %>%
  corrr::correlate() %>%
  mutate_if(is.numeric, funs(convert_dec2pct)) %>% 
  create_kable()
```

```{r viz_cor_dist, echo = FALSE, eval = FALSE, fig.height = 10, fig.width = 6}
# lab_title_cor_dist <- paste0("Correlation Matrix for RGB, HSL, and YUV\nEuclidean Distance Calculations")
# lab_subtitle_ord2_nba <-
#   stringr::str_replace_all(lab_subtitle_kclusts, lab_lgs, "NBA")
# viz_cor_dist <-
#   cor_dist_data %>% 
#   select_if(is.numeric) %>% 
#   GGally::ggcorr(label = TRUE, label_round = 2) +
#   viz_theme_base +
#   theme(legend.position = "right") +
#   viz_labs_base +
#   labs(title = lab_title_cor_dist, subtitle = lab_subtitle_ord2_nba)
# viz_cor_dist
```

The numbers indicate that there is some strong positive correlation, especially
between the RGB and YUV color schemas. This indicates that the conclusions
that I came to regarding most similar and dissimilar NBA team colors would not be
much different if using the HSL or YUV models instead of the RGB model.

### Distance vs. Similarity

To compare distance (Euclidean) with cosine similarity, I can create and use
a similar set of functions to those used for comparing distance methods.
To visualize the results in an interpretable manner,
I can use the `network_plot()` function from [Dr. Simon](https://drsimonj.svbtle.com/)'s 
[`corrr` package`](https://github.com/drsimonj/corrr).
This function is cool for visualizing correlation data in a way other than with a
traditional correlation matrix. [^fn_network_plot]

[^fn_network_plot]:
I could have actually used this same function to visualize the various distance methods
all in one plot.

```{r colors_ord2_nba_rgb_sim, include = FALSE}
do_pairwise_sim_byord <- function(data, ord) {
  data %>%
    filter(ord == ord) %>% 
    select(-ord) %>% 
    widyr::pairwise_similarity(name, rgb, value) %>% 
    rename(name1 = item1, name2 = item2) %>% 
    select(everything(), value = ncol(.)) %>% 
    mutate(ord = ord) %>% 
    select(ord, everything()) %>% 
    arrange(desc(value))
}

do_pairwise_sim <-
  function(data, ords = c("primary", "secondary")) {
    bind_rows(
      data %>%
        do_pairwise_sim_byord(ords[1]),
      data %>%
        do_pairwise_sim_byord(ords[2])
    )
  }
```

```{r corrr_rgb_hsl_dist_sim, echo = FALSE, fig.height = 6, fig.width = 10}
cor_sim_data <-
  colors_ord2_nba_rgb_tidy %>% 
  do_pairwise_sim() %>% 
  rename(rgb_sim = value) %>% 
  left_join(
    colors_ord2_nba_hsl_tidy %>% 
      do_pairwise_sim() %>% 
      rename(hsl_sim = value)
  ) %>% 
  left_join(
    colors_ord2_nba_yuv_tidy %>% 
      do_pairwise_sim() %>% 
      rename(yuv_sim = value)
  )
      
cor_distsim_data <-
  cor_dist_data %>%  
  left_join(cor_sim_data)
cor_rgbhslyuv_distsim <- 
  cor_distsim_data %>% 
  select_if(is.numeric) %>% 
  corrr::correlate()

lab_title_cor_rgbhslyuv_distsim <-
  paste0("Correlation Network for RGB, HSL, and YUV Euclidean Distance",
         "\nand Cosine Similarity Calculations")
lab_subtitle_ord2_nba <-
  stringr::str_replace_all(lab_subtitle_kclusts, lab_lgs, "NBA")

# NOTE: Get "'plot.tag' is not a valid theme element name error" if use `viz_theme_base`.
viz_cor_rgbhslyuv_distsim <-
  cor_rgbhslyuv_distsim %>% 
  corrr::network_plot(
    min_cor = 0, 
    colors = c("blue", "white", "red")
  ) +
  # teplot::theme_te() +
  # theme(
  #   legend.position = "bottom",
  #   # panel.background = element_rect(),
  #   panel.grid.major = element_blank(),
  #   panel.grid.minor = element_blank(),
  #   axis.ticks = element_blank(),
  #   plot.caption = element_text(face = "plain"),
  #   axis.text = element_blank()
  # ) +
  viz_theme_base +
  theme(legend.position = "bottom") +
  theme(axis.ticks = element_blank(), axis.text = element_blank()) +
  viz_labs_base +
  labs(title = lab_title_cor_rgbhslyuv_distsim, subtitle = lab_subtitle_ord2_nba)
viz_cor_rgbhslyuv_distsim
```

It's clear that the RGB and YUV schemas are fairly similar "within" both metrics--Euclidean distance
and cosine similarity--and both are relatively dissimilar to HSL. However, all three color models
show negative correlations "within" themselves when comparing the two metrics against one another.
(i.e. The RGB schema has a negative correlation when comparing its distance values
to its similarity values, and likewise for the HSL and YUV models.)

___So, which color model and which metric should be used?___
In my opinion, the RGB model seems like a good choice, both because it is
relatively similar to at least one other method (YUV) and because it is (probably) the most
relatable scheme to people who don't know much about color theory. For metric, I think that
the choice of Euclidean distance is valid. My Google search (which makes the case for YUV) makes the assumption
that Euclidean distance is being used. Additionally, a separate Google search
for "euclidean distance vs. cosine similarity" turns up 
[an easy-to-follow technical write-up](https://cmry.github.io/notes/euclidean-v-cosine)
that implies
that cosine similarity is probably not really appropriate for this kind of color analysis.

```{r nms_nfl, include = FALSE, eval = FALSE}
# [Maksim Horowitz](https://github.com/maksimhorowitz)
# [`nflscrapR` package](https://github.com/maksimhorowitz/nflscrapR)
# Need an alternative method of joining since the names are not
# identical across `teamcolors::teamcolors` and `nflscrapR::nflteams`.
identify_tm_mascot <- function(x) {
  stringr::str_replace_all(x, "^.*\\s", "")
}

nms_nfl <-
  teamcolors::teamcolors %>% 
  filter(league == "nfl") %>% 
  mutate(mascot = identify_tm_mascot(name)) %>% 
  inner_join(
    nflscrapR::nflteams %>%
      setNames(snakecase::to_snake_case(names(.))) %>%
      mutate_all(as.character) %>% 
      mutate(mascot = identify_tm_mascot(team_name)) %>% 
      select(slug = abbr, mascot),
    by = c("mascot")
  ) %>% 
  select(-mascot)


colors_tidy_ord2_nfl <-
  nms_nfl %>% 
  select(name, league, slug) %>% 
  inner_join(colors_tidy_ord2, by = c("name", "league"))
colors_ord2_nfl <-
  colors_tidy_ord2_nfl %>% 
  tidyr::spread(ord, hex)
```

## Conclusion

That's all I got for this topic. I hope that the techniques shown here are general enough
that they can be applied to any set of color to extract some fun (and meaningful) insight.


